---
title: "HW5_longitudinal_2"
author: "Jiseon Yang"
date: "2024-11-25"
output: word_document
---

# Data

```{r cars}

# Load necessary libraries
library(nlme)
library(glmmTMB)
library(reshape2)

# Load the data
strength_wide1 <- read.csv("C:/Users/jyang/OneDrive - Arizona State University/10 Classes_OneDrive/2024 8F_STP598_Logitudinal/zz HW/Longitudinal_HW3/strength.csv", header = FALSE)
names(strength_wide1) <- c("ID", "Group", "Week1", "Week2", "Week3", "Week4", "Week5", "Week6", "Week7")

# Convert data from wide format to long format
strength_long1 <- melt(strength_wide1, id.vars = c("ID", "Group"),
                       variable.name = "Week", value.name = "Strength")

# Convert variables to factors
strength_long1$Weekf <- as.factor(strength_long1$Week)
strength_long1$IDf <- as.factor(strength_long1$ID)
strength_long1$Groupf <- as.factor(strength_long1$Group)

# Check levels to prevent errors
if (length(levels(strength_long1$Groupf)) < 2) stop("Group must have at least two levels.")
if (length(levels(strength_long1$Weekf)) < 2) stop("Week must have at least two levels.")

# Convert `Week` to numeric after removing the "Week" prefix
strength_long1$Week_num <- as.numeric(gsub("Week", "", strength_long1$Weekf))

# Remove rows with missing values
strength_long1 <- na.omit(strength_long1)

```

# Question 2

## Use REML estimation and fit a linear mixed model for growth curves with random intercept and random linear slope. Include main effect for program, linear time trend, linear time by program interaction, and quadratic time trend as fixed effects.  

```{r}
# library(nlme)
# library(glmmTMB)
# library(reshape2)
# library(car)  # For Type III tests
# library(lmerTest)  # Extends lme4 for compatibility with Anova()
library(lme4)

# Fit the linear mixed model
fit <- lmer(
  Strength ~ Groupf + Week_num + I(Week_num^2) + Week_num:Groupf + 
    (1 + Week_num | IDf),  # Random intercept and slope for each subject
  # (Week_num + Week_num | IDf) produce the same result
  data = strength_long1,
  REML = TRUE
)

# Model summary
summary(fit)

# Perform Type III ANOVA
type3_anova <- anova(fit, type = 3)
print(type3_anova)

```

```{r}
library(ggplot2)

ggplot(strength_long1, aes(x = Week_num, y = Strength, color = Groupf, group = IDf)) +
  geom_line(alpha = 0.5) +
  labs(title = "Individual Growth Curves by Group",
       x = "Week",
       y = "Strength") +
  theme_minimal()


ggplot(strength_long1, aes(x = Week_num, y = Strength, color = Groupf)) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  labs(title = "Group-level Trends",
       x = "Week",
       y = "Average Strength") +
  theme_minimal()
```





# Question 6
```{r}
# Extract the random effects variance-covariance matrix D (for intercept and slope)
D <- matrix(c(9.2181, -0.18 * sqrt(9.2181 * 0.1255), 
              -0.18 * sqrt(9.2181 * 0.1255), 0.1255), nrow = 2)
D

# Define Z_i for 7 time points (1 column for intercept, 1 column for slope)
time_points <- 1:7
Z_i <- cbind(1, time_points)
Z_i

# Calculate Zi * D * Zi'
Zi_D_Zi_t <- Z_i %*% D %*% t(Z_i)
Zi_D_Zi_t

# Add the residual variance (Ri)
Ri <- diag(0.6006, nrow = 7, ncol = 7)
Ri

# covariance matrix Σi(α)
Sigma_i_alpha <- Zi_D_Zi_t + Ri

# Print 
Sigma_i_alpha
```


# Question 7
## 1. Estimating the Model with Random Intercept and Random Slope (using ML)
```{r}
# Fit model with random intercept and random slope using ML
model_random_intercept_slope_ml <- lmer(Strength ~ Groupf + Week_num + I(Week_num^2) + Week_num:Groupf + 
               (1 + Week_num | IDf), data = strength_long1, REML = FALSE)

# Summary of the model
summary(model_random_intercept_slope_ml)
anova(model_random_intercept_slope_ml)
```

## 2. Estimating the Model with Only Random Intercept (using ML)

```{r}
# Fit model with random intercept only using ML
model_random_intercept_ml <- lmer(Strength ~ Groupf + Week_num + I(Week_num^2) + Week_num:Groupf + 
                (1 | IDf), data = strength_long1, REML = FALSE)

# Summary of the model
summary(model_random_intercept_ml)
anova(model_random_intercept_ml)
```


## 3. Compare Models: Random Intercept and Random Slope vs. Random Intercept Only
```{r}
# Perform likelihood ratio test
anova(model_random_intercept_ml, model_random_intercept_slope_ml)
```

# Question 8

```{r}
# Refit the random intercept and random slope model using REML
model_random_intercept_slope_reml <- lmer(Strength ~ Groupf + Week_num + I(Week_num^2) + Week_num:Groupf + 
            (1 + Week_num | IDf), data = strength_long1, REML = TRUE)
summary(model_random_intercept_slope_reml)
anova(model_random_intercept_slope_reml)
```


# Question10-11
```{r}
# Fit the nested model without grand mean
nested_model <- 
  lmer(Strength ~ -1 + Groupf + Week_num:Groupf + I(Week_num^2):Groupf + 
                     (Week_num + Week_num | IDf), data = strength_long1, REML = TRUE) 
# (Week_num + Week_num | IDf) .. same result

summary(nested_model)
anova(nested_model)
```


# Question 13
```{r}
library(lme4)

# Fit the mixed-effects model with random intercept and slope
model <-
  lmer(Strength ~ -1 + Groupf + Week_num:Groupf + I(Week_num^2):Groupf +
                     (1 + Week_num | IDf), data = strength_long1, REML = TRUE) #(Week_num + Week_num | IDf) produce same result


# Obtain BLUP estimates for all subjects
blup_estimates <- ranef(model)$IDf

# Extract BLUP estimates for the first subject in the WI program
first_WI_subject <- strength_long1[strength_long1$Groupf == "WI", "IDf"][1]
blup_first_WI <- blup_estimates[first_WI_subject, ]
print(blup_first_WI)

# get adjusted_intercept and adjusted_slope
adjusted_intercept <- fixed_intercept_WI + subject_1_blup["(Intercept)"]
adjusted_slope <- fixed_slope_WI + subject_1_blup["Week_num"]
print(adjusted_intercept)
print(adjusted_slope )

# Filter data for the first subject in the WI program
first_subject_data <- subset(strength_long1, IDf == first_WI_subject)

# Fit an OLS model to the data of the first subject
ols_model <- lm(Strength ~ Week_num, data = first_subject_data)
ols_coefficients <- coef(ols_model)# Extract OLS coefficients
print(ols_coefficients)

# Step 5: Compare the estimates
cat("\nComparison of BLUP and OLS estimates for WI, Subject 1:\n")
comparison <- data.frame(
  Parameter = c("Intercept", "Slope"),
  BLUP = c(blup_first_WI["(Intercept)"], blup_first_WI["Week_num"]),
  OLS = c(ols_coefficients[1], ols_coefficients[2])
)
print(comparison)


```



```{r}
library(lme4)

# Fit the mixed-effects model with random intercept and slope
model <- lmer(Strength ~ -1 + Groupf + Week_num:Groupf + I(Week_num^2):Groupf +
                (1 + Week_num | IDf), data = strength_long1, REML = TRUE)

# Step 1: Extract fixed effects for the WI group
fixed_effects <- fixef(model)
fixed_intercept_WI <- fixed_effects["GroupfWI"]
fixed_slope_WI <- fixed_effects["GroupfWI:Week_num"]

# Step 2: Obtain BLUP estimates for all subjects
blup_estimates <- ranef(model)$IDf

# Step 3: Extract BLUP estimates for the first subject in the WI program
first_WI_subject <- strength_long1[strength_long1$Groupf == "WI", "IDf"][1]
blup_first_WI <- as.numeric(blup_estimates[first_WI_subject, ])
names(blup_first_WI) <- colnames(blup_estimates)  # Ensure column names match

cat("\nBLUP estimates for WI, Subject 1:\n")
print(blup_first_WI)

# Step 4: Calculate Adjusted BLUP Intercept and Slope
adjusted_intercept <- fixed_intercept_WI + blup_first_WI["(Intercept)"]
adjusted_slope <- fixed_slope_WI + blup_first_WI["Week_num"]
cat("\nAdjusted BLUP Intercept (WI, Subject 1):", adjusted_intercept, "\n")
cat("Adjusted BLUP Slope (WI, Subject 1):", adjusted_slope, "\n")

# Step 5: Fit an OLS model to the data of the first subject
first_subject_data <- subset(strength_long1, IDf == first_WI_subject)
ols_model <- lm(Strength ~ Week_num, data = first_subject_data)
ols_coefficients <- coef(ols_model)
cat("\nOLS estimates for WI, Subject 1:\n")
print(ols_coefficients)

# Step 6: Compare the BLUP Adjusted Intercept/Slope and OLS results
comparison <- data.frame(
  Parameter = c("Intercept", "Slope"),
  BLUP_Adjusted = c(adjusted_intercept, adjusted_slope),
  OLS = c(ols_coefficients[1], ols_coefficients[2])
)
cat("\nComparison of Adjusted BLUP and OLS estimates for WI, Subject 1:\n")
print(comparison)

```
The BLUP Adjusted Intercept and Slope provide a balanced estimate by combining the subject-specific variability and the population-level fixed effects. The differences arise because the BLUP accounts for both fixed effects and random effects (subject-specific variations), while the OLS is fitted solely to the data of Subject 1. The BLUP incorporates shrinkage towards the population mean, making it more robust in cases with fewer data points for individual subjects.


```{r}

```




