---
title: "HW2_JY"
output: word_document
date: "2023-09-04"
---

## 2.5a. Estimate the change in the mean service time when the number of copiers serviced increases by one. Use a 90 percent confidence interval. Interpret your confidence interval.

90% CI for β1 = b1 +/- t(1-α/2, df)*sb1
•	b1 = 15.0352 mins
•	α = 0.1
•	df = 45-2 = 43
•	qt(1-0.1/2, 43 ) = 1.681071
•	Sb1 = estm.std.err(b1) =root( Σe^2 /(n-p)) = 0.4831
•	LL <- b1 - 1.681071*sb1
•	UL <- b1 + 1.681071*sb1

### set work directory
```{r }
setwd("C:/Users/jyang/OneDrive - Arizona State University/10 Classes_OneDrive/2023_STP530_Regression")
```

### import data
```{r }
HW2.data <- read.table("CH01PR20.txt")
head(HW2.data)
colnames(HW2.data) <- c("min", "ser.no")
head(HW2.data)
```

### regression model
y.hat = b0 + b1X
```{r }
HW2.mod <- lm(min ~ ser.no, data = HW2.data)
summary(HW2.mod)
plot(min ~ ser.no, data = HW2.data, pch = 16, col = "darkgrey")
abline(coef(HW2.mod), col="red")
```

### 90% CI for b1

90% confidence interval (CI) for the expected changes in the mean service time for each 1-unit increase in the number of copiers serviced = 90% CI for β1.

##### 90% cI for b1 (by R code)
```{r }
confint(HW2.mod, level = .90)
```
##### 90% CI for b1 = b1 +/- t((1-a)/2,df)*sb1
15.0352 - qt(p=.95, df=43) * 0.4831 # lower limit
15.0352 + qt(p=.95, df=43) * 0.4831 # upper limit
```{r }
b1 <- 15.0352
df <- 43
qt(1-0.1/2,43)
sb1 <- 0.4831
LL <- b1 - 1.681071*sb1 
UL <- b1 + 1.681071*sb1 
LL
UL
print("The estimated mean service time changed by each increase in the number of copier serviced is 15.0352 mins. With 90% confidence, we estimate that the expected mean service time changes by somewhat between 14.22307 and 15.84733 for each one-unit increase in the No. of copier served")

```
==> Thus, the estimated mean service time changed by each increase in the number of copier serviced is 15.0352 mins. With 90% confidence, we estimate that the expected mean service time changes by somewhat between 14.22307 and 15.84733 for each one-unit increase in the No. of copier served.


## 2.5e. Does bo give any relevant information here about the "start-up" time on calls-Le., about the time required before service work is begun on the copiers at a customer location?
From above summary, 
•	b0 = 0.5802
•	sb0 = 2.8039
•	p.value.b0 = 0.837

==> Thus, the 'start-up' time on calls (the intercept) is not likely -0.5802 since it is a negative value. Also, it's not statistically significant with a p-value of 0.837.

## 5.14a. Obtain a 90 percent confidence interval for the mean gervice time on calls in which six copiers are serviced. Interpret your confidence interval.

# 95% confidence interval and prediction interval for X=6
```{r }
Y.hat.6 <- data.frame("ser.no"=6)
predict(HW2.mod, interval="confidence", newdata=Y.hat.6, level=.90) # confidence interval
predict(HW2.mod, interval="prediction", newdata=Y.hat.6, level=.90) # prediction interval  
```

# plot 95% confidence or prediction interval (CI, PI) for X=6
```{r }
plot(min ~ ser.no, data = HW2.data, pch = 16, col = "darkgrey",
     main = "HW2_Copier maintenance",
     xlab = "No. of copiers serviced",
     ylab = "Mean service time")
min_val <- min(HW2.data$ser.no, na.rm = TRUE)
max_val <- max(HW2.data$ser.no, na.rm = TRUE)
Y.hat.6 <- data.frame("ser.no" = seq(from = min_val, to = max_val, by = 1))
HW2.conf <- predict(HW2.mod, interval="confidence", newdata=Y.hat.6, level=.90)
head(HW2.conf,6)
lines(HW2.conf[, 1], lwd=2, col = "red")
lines(HW2.conf[, 2], lwd=2, lty = 3, col = "forestgreen" )
lines(HW2.conf[, 3], lwd=2, lty = 3, col = "forestgreen" )
HW2.pred <- predict(HW2.mod, interval="prediction", newdata=Y.hat.6, level=.90)
head(HW2.pred)
lines(HW2.pred[, 2], lwd=2, lty = 2, col = "blue" )
lines(HW2.pred[, 3], lwd=2, lty = 2, col = "blue" )
legend("topleft", legend = c("regression", "confidence", "prediction"),
       lty=c(1,3,2), lwd = 2, col=c("red", "forestgreen", "blue"), cex=1)
```

