---
title: "HW2"
author: "Jiseon Yang"
date: "2024-10-01"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.

```{r }
# Define the matrix
A <- matrix(c(1, 0, 0, 0,
              -1, 1, 0, 0,
              0, -1, 1, 0,
              0, 0, -1, 1,
              0, 0, 0, -1), 
            nrow = 5, ncol = 4, byrow = TRUE)

library(pracma)

# Gram-Schmidt process
Q <- gramSchmidt(A)$Q

# Output the orthonormal matrix
gramSchmidt(Q)
```

## 2 (a) Profile plot 
•	WI (Weights Increasing): The trend appears to be mostly linear but could have a slight quadratic component due to the accelerating increase in later weeks.
•	RI (Repetitions Increasing): A quadratic trend is likely, with an increase followed by a leveling off of strength.
•	CONT (Control): A flat or slightly negative linear trend, indicating no major polynomial trend.

```{r }
# Load necessary libraries
library(dplyr)
library(ggplot2)
#install.packages("tidyr")
library(tidyr)
library(reshape2)

# Load the data
strength_wide <- read.table("C:/Users/jyang/OneDrive - Arizona State University/10 Classes_OneDrive/2024 8F_STP598_Logitudinal/zz HW/Longitudinal_HW2/strength.dat", header = FALSE)

# Add proper column names for better understanding
names(strength_wide) <- c("ID", "Group", "Week1", "Week2", "Week3", "Week4", "Week5", "Week6", "Week7")

# Convert data from wide format to long format
strength_long <- melt(strength_wide, id.vars = c("ID", "Group"),
                      variable.name = "Week", value.name = "Strength")

# Convert 'Week' to a numeric variable (removing 'Week' prefix)
strength_long$Week <- as.numeric(gsub("Week", "", strength_long$Week))

length(unique(strength_long$ID))
table(strength_wide$Group)


# Generate the plot
strength_long %>% 
  group_by(Group, Week) %>% 
  summarise(Strength = mean(Strength, na.rm = TRUE), .groups = "drop") %>% 
  ggplot(aes(x = Week, y = Strength, color = as.factor(Group))) + 
  geom_point() +
  geom_line() +
  ylab("Strength Means") + 
  scale_color_discrete(name = "Group") +
  ggtitle("Profile Plot of Strength Over Time by Treatment Group") +
  xlab("Week")

```
## 2 (c) Fit the Split-Plot Model: ANOVA approach
-. use aov() function
-. best suit for banlanced model (not for unbalanced model)
-. contrast function 

1) aov.fit1: Split-plot model with a random effect for ID handled by the Error(IDf) term. Best suited for balanced designs.
2) aov.fit2: ANOVA model treating ID as a fixed effect. Overfits when there are many participants and treats individual variability as a fixed difference rather than random variability.
3) lm.fit3: Similar to aov.fit1, it treats ID as a fixed effect using a standard linear regression model. Again, this leads to overfitting when there are many participants.
4) lmer.fit4: Mixed-effects model that treats ID as a random effect. This is the best choice for repeated measures data or hierarchical data, and it can handle unbalanced data. It provides a flexible structure to account for both fixed and random effects, and it generalizes better than models treating ID as a fixed effect.
```{r }

# Load the data
strength_wide <- read.table("C:/Users/jyang/OneDrive - Arizona State University/10 Classes_OneDrive/2024 8F_STP598_Logitudinal/zz HW/Longitudinal_HW2/strength.dat", header = FALSE)

# Add proper column names for better understanding
names(strength_wide) <- c("ID", "Group", "Week1", "Week2", "Week3", "Week4", "Week5", "Week6", "Week7")

# Provide a Unique ID for each subject by combining ID and Group so that the same ID# is not considered to be the same subject accross different groups
strength_wide$UniqueID <- interaction(strength_wide$ID, strength_wide$Group)

# Convert data from wide format to long format (reshaping only once)
strength_long <- melt(strength_wide, id.vars = c("UniqueID", "Group"),
                      variable.name = "Week", value.name = "Strength")

# Remove 'Week' prefix and convert Week to a numeric variable
strength_long$Week <- as.numeric(gsub("Week", "", strength_long$Week))

# Convert Week, UniqueID, and Group to factors
strength_long$Weekf <- as.factor(strength_long$Week)
strength_long$UniqueIDf <- as.factor(strength_long$UniqueID)
strength_long$Groupf <- as.factor(strength_long$Group)

# Contrast
levels(strength_long$Weekf) # Check levels of Weekf
contrasts(strength_long$Weekf) <- contr.poly(7)

# 1) aov.fit1: Split-plot model with a random effect for ID handled by the Error(IDf) term. Best suited for balanced designs.
# aov.fit1 <- aov(Strength ~ Weekf + Groupf + Weekf*Group + Error(UniqueIDf), data = strength_long)
# aov.fit1
# summary(aov.fit1)

# 2) aov.fit2: ANOVA model treating ID as a fixed effect. 
aov.fit2 <- aov(Strength ~ Weekf + Groupf + Weekf*Group + UniqueIDf, data = strength_long)
summary(aov.fit2)

# 3) lm.fit3: Similar to aov.fit1, it treats ID as a fixed effect using a standard linear regression model. 
lm.fit3 <- lm(Strength ~ Weekf + Groupf + Weekf*Groupf + UniqueIDf, data = strength_long)
summary(lm.fit3)

# ANOVA
#anova(aov.fit1)
anova(aov.fit2)
anova(lm.fit3)
anova(lm.fit3)


```



## R Markdown



```{r }

```

## 2(a). Using New data table

```{r }
# Load necessary libraries
library(dplyr)
library(ggplot2)
#install.packages("tidyr")
library(tidyr)
library(reshape2)

# Load the data
strength_wide1 <- read.csv("C:/Users/jyang/OneDrive - Arizona State University/10 Classes_OneDrive/2024 8F_STP598_Logitudinal/zz HW/Longitudinal_HW2/strength.csv", header = FALSE)

str(strength_wide)

# Add proper column names for better understanding
names(strength_wide1) <- c("ID", "Group", "Week1", "Week2", "Week3", "Week4", "Week5", "Week6", "Week7")

head(strength_wide1)

# Convert data from wide format to long format
strength_long1 <- melt(strength_wide1, id.vars = c("ID", "Group"),
                      variable.name = "Week", value.name = "Strength")

# Convert 'Week' to a numeric variable (removing 'Week' prefix)
strength_long1$Week <- as.numeric(gsub("Week", "", strength_long1$Week))

length(unique(strength_long1$ID))
table(strength_wide1$Group)


# Generate the plot
strength_long1 %>% 
  group_by(Group, Week) %>% 
  summarise(Strength = mean(Strength, na.rm = TRUE), .groups = "drop") %>% 
  ggplot(aes(x = Week, y = Strength, color = as.factor(Group))) + 
  geom_point() +
  geom_line() +
  ylab("Strength Means") + 
  scale_color_discrete(name = "Group") +
  ggtitle("Profile Plot of Strength Over Time by Treatment Group-1") +
  xlab("Week")


```

## 2(c) Using New data table

```{r }

# Convert data from wide format to long format (reshaping only once)
strength_long1 <- melt(strength_wide1, id.vars = c("ID", "Group"),
                      variable.name = "Week", value.name = "Strength")

# Remove 'Week' prefix and convert Week to a numeric variable
strength_long1$Week <- as.numeric(gsub("Week", "", strength_long1$Week))

# Convert Week, ID, and Group to factors
strength_long1$Weekf <- as.factor(strength_long1$Week)
strength_long1$IDf <- as.factor(strength_long1$ID)
strength_long1$Groupf <- as.factor(strength_long1$Group)

# Contrast
levels(strength_long1$Weekf) # Check levels of Weekf
contrasts(strength_long1$Weekf) <- contr.poly(7)
contrasts(strength_long1$Weekf)

# 1) aov.fit5: Split-plot model with a random effect for ID handled by the Error(IDf) term. Best suited for balanced designs.
aov.fit5 <- aov(Strength ~ Weekf + Groupf + Weekf*Group + Error(IDf), data = strength_long1)
summary(aov.fit5)

# 2) aov.fit6: ANOVA model treating ID as a fixed effect. 
aov.fit6 <- aov(Strength ~ Weekf + Groupf + Weekf*Group + IDf, data = strength_long1)
summary(aov.fit6)

# 3) lm.fit7: Similar to aov.fit1, it treats ID as a fixed effect using a standard linear regression model. 
lm.fit7 <- lm(Strength ~ Weekf + Groupf + Weekf*Groupf + IDf, data = strength_long1)
summary(lm.fit7)

# 4) lmer.fit8: Mixed-effects model that treats ID as a random effect. Best recommended for repeated measures data. Also, it can handle unbalanced data. It provides a flexible structure to account for both fixed and random effects, and it generalizes better than models treating ID as a fixed effect.
# lmer.fit8 <- lmer(Strength ~ Weekf + Group + Weekf*Group + (1 | IDf), data = strength_long1)
# summary(lmer.fit8)

# ANOVA
#anova(aov.fit5)
anova(aov.fit6)
anova(lm.fit7)
# anova(lmer.fit8)

```

## 3. Extra credit 

```{r }

summary(lm.fit7)$coefficients


```


```{r }



```

