---
title: "HW4_3.25"
output: word_document
date: "2023-09-19"
---



### Import data
```{r}
# Clean up the workspace for the new analysis
rm(list=ls()) 

# Set the following to YOUR OWN folder
setwd("C:/Users/jyang/OneDrive - Arizona State University/10 Classes_OneDrive/2023_STP530_Regression")

# Import the dataset. The txt data file needs to exist in the folder above.
mydata <- read.table("APPENC02.txt")
head(mydata)

# Rename the columns
colnames(mydata) <- c("ID","Country", "State", "Area.2mile", "Population", "Age18to34.pst", 
                      "Age65over.pst", "Physician.no", "bed.no", "crime.no", "hischool.pst",
                      "BS.pst", "Low.income.pst", "unemployment.pst", "Per.income.dollar",
                      "TO.income.m", "Geo"
                      ) # Rename the columns
head(mydata)

```

### Fitted plot, linear regression model and possible outliers
```{r}
# Fit the linear regression model
Y <- mydata$Physician.no

m1 <- lm(Y~ Population, data=mydata) #related to total population
plot(Y~ Population, data=mydata,
     xlab="Total polulation",
     ylab="The number of active physicians in a CDI")
abline(coef(m1), col="red", lty=3, lwd=2)
summary(m1)

m2 <- lm(Y~ bed.no, data=mydata) #related to number of hospital beds
plot(Y~ bed.no, data=mydata,
     xlab="Hospital beds")
abline(coef(m2), col="blue", lty=2, lwd=2)
summary(m2)

m3 <- lm(Y~ TO.income.m, data=mydata) #related to total personal income.
plot(Y~ TO.income.m, data=mydata,
     xlab="Total personal income milions")
abline(coef(m3), col="forestgreen",lty=1, lwd=2)
summary(m3)

# Possible outliners, by examining the histogram of each variable 
hist(mydata$Population)
hist(mydata$bed.no)
hist(mydata$TO.income.m)
# appears that there are a couple outliers
```


### Fit a multiple regression model.
```{r}
# scatter plot matrix 
pairs(mydata[, 5:16])

# regression models
mod1 <- lm(Y ~ Population, data=mydata, na.action=na.exclude)
summary(mod1)

mod2 <- lm(Y ~ bed.no, data=mydata, na.action=na.exclude)
summary(mod2)
 
mod3 <- lm(Y ~ TO.income.m, data=mydata, na.action=na.exclude)
summary(mod3)
```


# Problems:
###(1) Plot the residuals against the predictor (or Y-hat).
(a) Whether there seems to be a linear or non-linear relationship?
(b) Is the residuals heteroskedastic? variance (vertical spread) across the board
Funnel shape opening to the right: heteroskedasticity pattern, a Poisson distribution
```{r}
# Residuals vs. Predictor for Model 1 (Population)
plot(fitted(mod1), residuals(mod1), main="Residuals vs. Population")
abline(h=0, col="red")

# Residuals vs. Predictor for Model 2 (Number of Beds)
plot(fitted(mod2), residuals(mod2), main="Residuals vs. Number of Beds")
abline(h=0, col="blue")

# Residuals vs. Predictor for Model 3 (Total Personal Income)
plot(fitted(mod3), residuals(mod3), main="Residuals vs. Total Personal Income")
abline(h=0, col="forestgreen")


# Also, may plot the residuals against each X to see whether the nonlinear trend in the residuals might be potentially related to one of the Xs.
plot(mydata$Population, residuals(mod1))
plot(mydata$bed.no, residuals(mod2))
plot(mydata$TO.income.m, residuals(mod3))
```

### Regression outliers
```{r}
# Studentized residual plot
plot(fitted(mod1), rstudent(mod1))
plot(fitted(mod2), rstudent(mod2))
plot(fitted(mod3), rstudent(mod3))

# find out which point is the outlier
plot(fitted(mod1), rstudent(mod1), type="n")
text(fitted(mod1), rstudent(mod1), names(rstudent(mod1)))

plot(fitted(mod2), rstudent(mod2), type="n")
text(fitted(mod2), rstudent(mod2), names(rstudent(mod2)))

plot(fitted(mod3), rstudent(mod3), type="n")
text(fitted(mod3), rstudent(mod3), names(rstudent(mod3)))

```

# Problems:
### (2) Whether the residuals are normally distributed
Histogram or QQ-plot
QQ-plot: The more straight line, the more normal
Data do not fit well.
```{r}
# Histogram
hist(residuals(mod1), main="Histogram for Model 1 Residuals")
hist(residuals(mod2), main="Histogram for Model 2 Residuals")
hist(residuals(mod3), main="Histogram for Model 3 Residuals")

# Q-Q plot for model 1
qqnorm(residuals(mod1), main ="QQ-plot for Model 1 Residuals")
qqline(residuals(mod1), col="red", lwd=2)

# Q-Q plot for model 2
qqnorm(residuals(mod2), main ="QQ-plot for Model 2 Residuals")
qqline(residuals(mod2), col="blue", lwd=2)

# Q-Q plot for model 3
qqnorm(residuals(mod3), main ="QQ-plot for Model 3 Residuals")
qqline(residuals(mod3), col="forestgreen", lwd=2)
```

### Log-regression models 

```{r}
#-------------------------------------------------
# Transform Y with log(), fit the new model, and repeat the diagnostics
mydata$log.Physician.no <- log(mydata$Physician.no)
logmod1 <- lm(log.Physician.no ~ Population, data=mydata, na.action=na.exclude)
plot(mydata$Population, residuals(logmod1))
qqnorm(residuals(logmod1), main="QQ-plot for log.Mod-1 Residuals.")
qqline(residuals(logmod1), col="red", lwd=2)
summary(logmod1)

logmod2 <- lm(log.Physician.no ~ bed.no, data=mydata, na.action=na.exclude)
plot(mydata$bed.no, residuals(logmod2))
qqnorm(residuals(logmod2), main="QQ-plot for log.Mod-2 Residuals.")
qqline(residuals(logmod2), col="blue", lwd=2)
summary(logmod2)

logmod3 <- lm(log.Physician.no ~ TO.income.m, data=mydata, na.action=na.exclude)
plot(mydata$TO.income.m, residuals(logmod3))
qqnorm(residuals(logmod3), main="QQ-plot for log.Mod-3 Residuals.")
qqline(residuals(logmod3), col="blue", lwd=2)
summary(logmod3)
```

### Identify outliers
```{r}
# You can use the following code to identify the outlier
plot(fitted(logmod1), rstudent(logmod1), type='n')
text(fitted(logmod1), rstudent(logmod1), names(rstudent(logmod1)))
rstudent(logmod1)[1] # outlier check
outliers_gt_2 <- rstudent(logmod1)[abs(rstudent(logmod1)) > 2] # outliers >2
print(outliers_gt_2)

plot(fitted(logmod2), rstudent(logmod2), type='n')
text(fitted(logmod2), rstudent(logmod2), names(rstudent(logmod2)))
rstudent(logmod2)[1] # outlier check
outliers_gt_2 <- rstudent(logmod2)[abs(rstudent(logmod2)) > 2] # outliers >2
print(outliers_gt_2)

plot(fitted(logmod3), rstudent(logmod3), type='n')
text(fitted(logmod3), rstudent(logmod3), names(rstudent(logmod3)))
rstudent(logmod3)[1] # outlier check
outliers_gt_2 <- rstudent(logmod3)[abs(rstudent(logmod3)) > 2] # outliers >2
print(outliers_gt_2)

# Remove the outliers discovered and call the new data as XX.reduced
mydata.reduced <- mydata[-c(1, 2),] #remove two outliers
```

### Log-regression models on the 'reduced' data
```{r}
# Fit the log-transformed model on the reduced data
logmod1.reduced <- lm(log(Physician.no) ~ Population, data=mydata.reduced, na.action=na.exclude)
logmod2.reduced <- lm(log(Physician.no) ~ bed.no, data=mydata.reduced, na.action=na.exclude)
logmod3.reduced <- lm(log(Physician.no) ~ TO.income.m, data=mydata.reduced, na.action=na.exclude)

# Residual plot of the log-transformed model
plot(fitted(logmod1.reduced), rstudent(logmod1.reduced), main="log.mod1.reduced")
plot(fitted(logmod2.reduced), rstudent(logmod2.reduced), main="log.mod2.reduced")
plot(fitted(logmod3.reduced), rstudent(logmod3.reduced), main="log.mod3.reduced")
```

### Q-Q plot of the log-transformed model
It looks better now.
```{r}
# Q-Q plot of the log-transformed model
qqnorm(residuals(logmod1.reduced), main="QQ-plot for log.Mod1.reduced Residuals")
qqline(residuals(logmod1.reduced), col="red", lwd=2)

qqnorm(residuals(logmod2.reduced), main="QQ-plot for log.Mod2.reduced Residuals")
qqline(residuals(logmod2.reduced), col="blue", lwd=2)

qqnorm(residuals(logmod3.reduced), main="QQ-plot for log.Mod3.reduced Residuals")
qqline(residuals(logmod3.reduced), col="forestgreen", lwd=2)
```

### summary table
```{r}
summary(logmod1.reduced)
summary(logmod2.reduced)
summary(logmod3.reduced)
```


