---
title: "Final_Longitudinal_Short"
author: "Jiseon Yang"
date: "2024-11-23"
output: word_document
---




```{r setup, include=FALSE}
options(width = 180)  # Set the desired width (e.g., 180 characters)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tibble.print_max = 300, tibble.print_min = 300) # Adjust numbers as needed

```


### New Data
```{r }
# Load necessary libraries
library(tidyr)
library(dplyr)
library(tidyverse)
library(reshape2)

# Load data: ISS_CFU_size_new.csv
ISS_CFU_size_new <- read.csv("C:/Users/jyang/ASU Dropbox/Jiseon Yang/2 PROJJECT/2019_NASA CoEvolve/10 Micro17-Ground investigation/ISS_CFU_size_new.csv")

# Reshape the data into long format
data_long <- ISS_CFU_size_new %>%
  pivot_longer(
    cols = c("Ral_L", "Ral_s", "Sph_L", "Sph_s"),
    names_to = c("Species", "Size"),
    names_sep = "_",
    values_to = "CFU"
  )

# Column type
data_long <- data_long %>%
  mutate(
    Days = as.numeric(Days),
    Daysf = as.factor(Days),
    ID = as.factor(ID),
    trt = as.factor(trt) # 'trt' column already in your data
  )
# Log10_CFU
data_long$log_CFU <- log10(data_long$CFU)
# View the reshaped data
head(data_long)

# check NA
cat("\n NA = ",sum(is.na(data_long)))
cat("\n Zero CFU = ", sum(data_long$CFU == 0, na.rm = TRUE))
data_long <- na.omit(data_long)
cat("\n NA = ",sum(is.na(data_long)))


# Aggregate CFU for each species (L + S)
species_totals <- data_long %>%
  group_by(ID, Days, Daysf, trt, Species) %>%
  summarize(Species_CFU = sum(CFU, na.rm = TRUE))

# Log10_Species_CFU
species_totals$log_Species_CFU <- log10(species_totals$Species_CFU)
head(species_totals)

cat("\n NA = ",sum(is.na(species_totals)))
cat("\n Zero CFU = ", sum(species_totals$Species_CFU == 0, na.rm = TRUE))

```



# spaghetti plot and profile plot 
```{r}

library(ggplot2)

# Create spaghetti plot with modified scales
ggplot(data = species_totals, aes(x = Days, y = log_Species_CFU, color = Species, group = interaction(ID, Species))) +
  geom_line(alpha = 0.7) +  # Lines connecting data points for each individual
  geom_point(alpha = 0.6, size = 1.5) + # Points for observed data
  facet_wrap(~ trt, ncol = 1, scales = "free_y") + # Separate panels for each treatment group
  labs(title = "A. Individual Spaghetti Plots of Log Species CFU Over Days",
       x = "Days",
       y = "Log Species CFU",
       color = "Species") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c("Ral" = "blue", "Sph" = "darkorange")) + # Customize colors for species
  scale_y_continuous(limits = c(6.5, 9.5), breaks = seq(6.5, 9.5, 0.5)) + # Set y-axis scale
  scale_x_continuous(breaks = seq(0, max(species_totals$Days), 4)) # Set x-axis scale with interval 7

```


```{r}
library(ggplot2)
ggplot(species_totals, aes(x = Days, y = log_Species_CFU, color = trt, linetype = Species)) +
  geom_line(stat = "summary", fun = mean) +
  labs(title = "B. Growth Patterns Over Time", y = "Species population (log-CFU)", x = "Days") +
  theme(legend.position = "bottom") +
  #scale_y_continuous(limits = c(6.5, 9.5), breaks = seq(6.5, 9.5)) + # Set y-axis scale
  scale_x_continuous(breaks = seq(0, max(species_totals$Days), 10)) # Set x-axis scale with interval 7

```


## Research Question 1:
Do treatment conditions (LS vs. 1G) significantly alter the population dynamics of Ralstonia (Ral) and Sphingomonas (Sph) over time?
### log_Species_CFU ∼ Days * trt * Species +( 1∣ID)
#### Linear mixed model fit by REML
```{r}
# Linear mixed model fit by REML

library(lme4)

library(glmmTMB)

# Using lmer (random effects)
model_mixed <- lmer(log_Species_CFU ~ Days * trt * Species + (1 | ID), data = species_totals, REML = TRUE)
summary(model_mixed)
cat("\n")


library(lmerTest)
library(car)


anova(model_mixed, type = "III")
cat("\n *AIC(model_mixed) = ",AIC(model_mixed)) #  1487.367

```


```{r}
# Using gls (with AR(1) correlation)
library(nlme)
model_ar1 <- gls(
  log_Species_CFU ~ Days * trt * Species,
  correlation = corAR1(form = ~ 1 | ID),
  data = species_totals,
  method = "REML"
)
summary(model_ar1)
anova(model_ar1)
cat("\n *AIC(model_ar1) = ",AIC(model_ar1)) #  1152.575
AIC(model_mixed, model_ar1)


# Diagnostic checks
library(performance)

#check_model(model1_AR1)
check_model(model_ar1)

```


#Research Question 2:
What is the best-fit temporal autocorrelation structure?

### Days_scaled
```{r}
# Days_scaled
species_totals$Days_log <- log(species_totals$Days+1) # natural log
head(species_totals)
```
###(a) Unstructured (UN) Covariance -- fail to run
This allows each pair of repeated measures to have its own covariance.
```{r}

any(is.na(species_totals))

#install.packages("nlme")
library(nlme)

# model_UN <- gls(
#   log_Species_CFU ~  Days_log:Species+ trt:Species,
#   data = species_totals,
#   correlation = corSymm(form = ~ 1 | ID),
#   method = "REML"
# )
# summary(model_UN)

```

###b) Compound Symmetry (CS) Covariance
Assumes constant variance and equal correlation between all pairs of time points.
```{r}
model_CS <- gls(
  log_Species_CFU ~ Days_log * trt * Species,
  data = species_totals,
  correlation = corCompSymm(form = ~ Days_log | ID),
  method = "REML"
)
summary(model_CS)


model_cs <- gls(
  log_Species_CFU ~ Days * trt * Species,
  data = species_totals,
  correlation = corCompSymm(form = ~ Days | ID),
  method = "REML"
)
summary(model_cs)

# Diagnostic checks
library(performance)
check_model(model_CS)
check_model(model_cs)
```

###(c) Autoregressive (AR(1)) Covariance
Assumes that correlations decay exponentially with increasing time lag.
The correlation structure is applied within each group identified by ID.
The time points for correlation decay are assumed to be equally spaced and do not require an explicit time variable.
```{r}
# 1.	How do species, treatment, and time independently and interactively affect CFU counts? Are there specific combinations of time, species, and treatment that show unique trends? (The most comprehensive model for evaluating time, treatment, and species effects, along with all possible interactions.) {log_Species_CFU ~ trt * Species * Daysf} --------Not working Singular
model1_AR1 <- gls(
  log_Species_CFU ~ trt * Species * Days_log, #{Daysf:Species:trt} -- dont work
  data = species_totals,
  correlation = corAR1(form = ~ 1 | ID),
  method = "REML"
)
summary(model1_AR1)

model_ar1 <- gls(
  log_Species_CFU ~ trt * Species * Days, #{Daysf:Species:trt} -- dont work
  data = species_totals,
  correlation = corAR1(form = ~ 1 | ID),
  method = "REML"
)
summary(model_ar1)

anova(model1_AR1,model_ar1)

# Diagnostic checks
library(performance)
check_model(model1_AR1)
check_model(model_ar1)

```

###(d) Toeplitz (TOEP) Covariance
Allows correlations to vary with time lag but remains constant for observations at the same lag.
```{r fig.height=8, fig.width=8 }
library(glmmTMB)


model1_toep <- glmmTMB(
  log_Species_CFU ~ Days_log*Species*trt + toep(Days_log + 0 | ID),
  data = species_totals,
  #dispformula = ~ 0, # Assume Residual variance is the same across trt groups.
  #dispformula = ~ species, # Residual variance is allowed to differ by species.
  #Without dispformula # Residual variance is estimated freely for each group.
  REML = TRUE)

summary(model1_toep) # 1487.2

############# select model2_toep
model2_toep <- glmmTMB(
  log_Species_CFU ~ Days_log*Species*trt + toep(Days_log + 0 | ID),
  data = species_totals,
  #dispformula = ~ 0, # Assume Residual variance is the same across trt groups.
  dispformula = ~ Species, # Residual variance is allowed to differ by species.
  #none  # Residual variance is estimated freely for each group.
  REML = TRUE)

summary(model2_toep) # 1463.8

# Diagnostic checks
library(performance)
check_model(model1_toep)
check_model(model2_toep)

```

### (e) sum
```{r}
# Call AIC, BIC, and logLik for each model manually
models <- list(
  CS = model_CS,
  cs = model_cs,
  AR1 = model1_AR1,
  ar1 = model_ar1,
  TOEP = model1_toep,
  TOEP_disp_species = model2_toep
  #TOEP_scaled = model_toep_scaled
)

# Create an empty data frame to store the results
results <- data.frame(
  Model = character(),
  AIC = numeric(),
  BIC = numeric(),
  logLik = numeric(),
  stringsAsFactors = FALSE
)

# Loop through each model and extract the metrics
for (model_name in names(models)) {
  model <- models[[model_name]]
  
  if (!is.null(model)) {
    results <- rbind(results, data.frame(
      Model = model_name,
      AIC = AIC(model),
      BIC = BIC(model),
      logLik = logLik(model)
    ))
  }
}

# Display the results
print(results)
```
\n CONCLUSION:
\n The AR(1) model, with its lowest AIC and BIC values, highest log-likelihood, and parsimonious structure, provides the best fit for modeling the CFU dynamics, effectively capturing the exponential decay in correlations over time while maintaining computational efficiency.


#Research Question 3:
What are the differences in growth trajectories between Ralstonia and Sphingomonas under LS and 1G conditions over time?

#### LMM_poly
```{r}
model_species_REML_poly2 <- lmer(
  log_Species_CFU ~ poly(Days, 2) * trt * Species + (1 | ID),
  data = species_totals, REML = TRUE)

model_species_REML_poly3 <- lmer(
  log_Species_CFU ~ poly(Days, 3) * trt * Species + (1 | ID),
  data = species_totals, REML = TRUE)

model_species_REML_poly4 <- lmer(
  log_Species_CFU ~ poly(Days, 4) * trt * Species + (1 | ID),
  data = species_totals, REML = TRUE)

model_species_REML_poly5 <- lmer(
  log_Species_CFU ~ poly(Days, 5) * trt * Species + (1 | ID),
  data = species_totals, REML = TRUE)

model_species_REML_poly6 <- lmer(
  log_Species_CFU ~ poly(Days, 6) * trt * Species + (1 | ID),
  data = species_totals, REML = TRUE)

model_species_REML_poly8 <- lmer(
  log_Species_CFU ~ poly(Days, 8) * trt * Species + (1 | ID),
  data = species_totals, REML = TRUE)

model_species_REML_poly10 <- lmer(
  log_Species_CFU ~ poly(Days, 10) * trt * Species + (1 | ID),
  data = species_totals, REML = TRUE)

summary(model_species_REML_poly10)
anova(model_species_REML_poly10, type = "III")

# compare
anova(model_mixed, model_species_REML_poly2) # p < 2.2e-16
anova(model_species_REML_poly5, model_species_REML_poly6) # p < 2.2e-16
anova(model_species_REML_poly6, model_species_REML_poly10) # p < 2.2e-16

cat("\n *AIC(model_species_REML) = ",AIC(model_mixed)) # 1487.367
cat("\n *AIC(model_species_REML_poly2) = ",AIC(model_species_REML_poly2))# 1350.321
cat("\n *AIC(model_species_REML_poly3) = ",AIC(model_species_REML_poly3))# 1229.857
cat("\n *AIC(model_species_REML_poly4) = ",AIC(model_species_REML_poly4))# 1032.717
cat("\n *AIC(model_species_REML_poly5) = ",AIC(model_species_REML_poly5))# 911.3526
cat("\n *AIC(model_species_REML_poly6) = ",AIC(model_species_REML_poly6))# 783.9919
cat("\n *AIC(model_species_REML_poly8) = ",AIC(model_species_REML_poly8))# 502.4344
cat("\n *AIC(model_species_REML_poly10) = ",AIC(model_species_REML_poly10)) #274.3817
```

#### ar1_poly_nested

## 1. Polynomial Terms for AR(1) Structure with Nested Random Effects (Days_log)
```{r}
# model_species_REML_poly10

# LMM_poly_ML
LMM_poly_ML <- lmer(
  log_Species_CFU ~ poly (Days, 10) * trt * Species + (1 | ID),
  data = species_totals,
  REML = FALSE
)


############ AR(1)_poly_nested (Days-log)


# Add log-polynomial terms to species_totals
species_totals <- species_totals %>%
  mutate(
    Days_log_sq = Days_log^2,
    Days_log_cub = Days_log^3,
    Days_log_4th = Days_log^4,
    Days_log_5th = Days_log^5,
    Days_log_6th = Days_log^6,
    Days_log_7th = Days_log^7,
    Days_log_8th = Days_log^8,
    Days_log_9th = Days_log^9,
    Days_log_10th = Days_log^10
  )


#log-polynomial ar1_poly_nested
ar1_poly_nested <- gls(log_Species_CFU ~ 
                     (Days_log + Days_log_sq + Days_log_cub + Days_log_4th + Days_log_5th + Days_log_6th + Days_log_7th + Days_log_8th + Days_log_9th + Days_log_10th) : trt : Species,
                 data = species_totals, 
                 correlation = corAR1(form = ~ 1 | ID), 
                 method = "REML")
summary(ar1_poly_nested)

library(car)

# Type III ANOVA to test interactions
anova_results <- Anova(ar1_poly_nested, type = "III")
print(anova_results)


#log-polynomial ar1_poly_nested
ar1_poly_nested2 <- gls(log_Species_CFU ~ 
                     (Days_log + Days_log_sq + Days_log_cub + Days_log_4th + Days_log_5th + Days_log_6th + Days_log_7th + Days_log_8th + Days_log_9th + Days_log_10th) : Species : trt + trt:Species + trt + Species,
                 data = species_totals, 
                 correlation = corAR1(form = ~ 1 | ID), 
                 method = "REML")
summary(ar1_poly_nested2)
print(Anova(ar1_poly_nested2, type = "III"))


##################### Visualization new_data$predicted &2
# Visualization
library(ggplot2)


# Generate new data for prediction
new_data <- expand.grid(
  Days = seq(min(species_totals$Days), max(species_totals$Days), length.out = 100),
  trt = unique(species_totals$trt),
  Species = unique(species_totals$Species)
)

# Add transformed Days (log) and polynomial terms to the new data
new_data <- new_data %>%
  mutate(
    Days_log = log(Days),
    Days_log_sq = Days_log^2,
    Days_log_cub = Days_log^3,
    Days_log_4th = Days_log^4,
    Days_log_5th = Days_log^5,
    Days_log_6th = Days_log^6,
    Days_log_7th = Days_log^7,
    Days_log_8th = Days_log^8,
    Days_log_9th = Days_log^9,
    Days_log_10th = Days_log^10
  )

# Predict values using the best model (e.g., ar1_poly_nested or ar1_poly_nested2)
new_data$predicted <- predict(ar1_poly_nested, newdata = new_data)

# Plot predictions
ggplot(new_data, aes(x = Days_log, y = predicted, color = Species, linetype = trt)) +
  geom_line(size = 1) +
  labs(title = "Predicted CFU Over Time by Treatment and Species",
       y = "Log CFU",
       x = "Days_log") +
  scale_x_continuous(breaks = seq(0, max(species_totals$Days), 0.2))+
  theme_minimal() +
  theme(legend.position = "right")

new_data$predicted2 <- predict(ar1_poly_nested2, newdata = new_data)

# Plot predictions
ggplot(new_data, aes(x = Days_log, y = predicted2, color = Species, linetype = trt)) +
  geom_line(size = 1) +
  labs(title = "Predicted CFU Over Time by Treatment and Species (ar1_poly_nested2)",
       y = "Log CFU",
       x = "Days_log") +
  scale_x_continuous(breaks = seq(0, max(species_totals$Days), 0.2))+
  theme_minimal() +
  theme(legend.position = "right")


###################### AR(1)_poly_nested

# Add polynomial terms to species_totals
species_totals <- species_totals %>%
  mutate(
    Days_sq = Days^2,
    Days_cub = Days^3,
    Days_4th = Days^4,
    Days_5th = Days^5,
    Days_6th = Days^6,
    Days_7th = Days^7,
    Days_8th = Days^8,
    Days_9th = Days^9,
    Days_10th = Days^10
  )

#ar1_poly_nested3
ar1_poly_nested3 <- gls(log_Species_CFU ~ 
                     (Days + Days_sq + Days_cub + Days_4th + Days_5th + Days_6th + Days_7th + Days_8th + Days_9th + Days_10th):trt:Species,
                 data = species_totals, 
                 correlation = corAR1(form = ~ 1 | ID), 
                 method = "REML")
summary(ar1_poly_nested3)


# Type III ANOVA 
library(car)
anova_results <- Anova(ar1_poly_nested3, type = "III")
print(anova_results)


#ar1_poly_nested4
ar1_poly_nested4 <- gls(log_Species_CFU ~ 
                     (Days + Days_sq + Days_cub + Days_4th + Days_5th + Days_6th + Days_7th + Days_8th + Days_9th + Days_10th): Species : trt + trt:Species + trt + Species,
                 data = species_totals, 
                 correlation = corAR1(form = ~ 1 | ID), 
                 method = "REML")
summary(ar1_poly_nested4)
print(Anova(ar1_poly_nested4, type = "III"))


###################### anova and diagnostics

AIC( model_species_REML_poly10, LMM_poly_ML, ar1_poly_nested, ar1_poly_nested2, ar1_poly_nested3, ar1_poly_nested4)

check_model(LMM_poly_ML)
check_model(ar1_poly_nested)
check_model(ar1_poly_nested2)
check_model(ar1_poly_nested3)
check_model(ar1_poly_nested4)

#####################  Visualization new_data$predicted3 &4
# Visualization
library(ggplot2)


# Generate new data for prediction
new_data <- expand.grid(
  Days = seq(min(species_totals$Days), max(species_totals$Days), length.out = 100),
  trt = unique(species_totals$trt),
  Species = unique(species_totals$Species)
)

# Add transformed Days (log) and polynomial terms to the new data
new_data <- new_data %>%
  mutate(
    Days = Days,
    Days_sq = Days^2,
    Days_cub = Days^3,
    Days_4th = Days^4,
    Days_5th = Days^5,
    Days_6th = Days^6,
    Days_7th = Days^7,
    Days_8th = Days^8,
    Days_9th = Days^9,
    Days_10th = Days^10
  )

# Predict values using the best model (e.g., ar1_poly_nested or ar1_poly_nested2)
new_data$predicted3 <- predict(ar1_poly_nested3, newdata = new_data)

# Plot predictions
ggplot(new_data, aes(x = Days, y = predicted3, color = Species, linetype = trt)) +
  geom_line(size = 1) +
  labs(title = "Predicted CFU Over Time by Treatment and Species",
       y = "Log CFU",
       x = "Days") +
  scale_x_continuous(breaks = seq(0, max(species_totals$Days), 4))+
  theme_minimal() +
  theme(legend.position = "right")

new_data$predicted4 <- predict(ar1_poly_nested4, newdata = new_data)

# Plot predictions
ggplot(new_data, aes(x = Days, y = predicted4, color = Species, linetype = trt)) +
  geom_line(size = 1) +
  labs(title = "Predicted CFU Over Time by Treatment and Species",
       y = "Log CFU",
       x = "Days") +
  scale_x_continuous(breaks = seq(0, max(species_totals$Days), 4))+
  theme_minimal() +
  theme(legend.position = "right")

###############

anova(ar1_poly_nested,ar1_poly_nested2, ar1_poly_nested3,ar1_poly_nested4)

VarCorr(LMM_poly_ML)
```
```{r}
summary(ar1_poly_nested2)
 
print(Anova(ar1_poly_nested2, type = "III"))


```



## Introduce breakpoint and analyze data
### (a) Estimate breakpoint with linear regression.
```{r}

#install.packages("segmented")
library(segmented)

################### 1. Linear regression, find breakpoint: 
###################      :estimated breakpoint is Day 2

## Fit a Piecewise Regression Model
lm_initial <- lm(log_Species_CFU ~ Days * Species *trt, data = species_totals)
library(segmented)

# Add a breakpoint to the initial linear model
segmented_model <- segmented(lm_initial, seg.Z = ~ Days, psi = 8)  # Set an initial guess for the breakpoint (e.g., psi = 8)

summary(segmented_model)


# Extract the estimated breakpoint
breakpoint <- segmented_model$psi[2]
print(breakpoint)

##Visualization: Plot the Piecewise Regression
library(ggplot2)

# Create predicted values for visualization
species_totals$Predicted <- predict(segmented_model)

# Plot the observed data with piecewise regression
ggplot(species_totals, aes(x = Days, y = log_Species_CFU)) +
  geom_point(alpha = 0.6) +  # Observed data
  geom_line(aes(y = Predicted), color = "blue", size = 1) +  # Fitted piecewise regression
  labs(title = "Piecewise Regression with Breakpoint",
       x = "Days",
       y = "Log Species CFU") +
  scale_x_continuous(breaks = seq(0, max(species_totals$Days), 4))+
  theme_minimal()
```

### (b) Modeling with a Manually Set Breakpoint at Day 8.
```{r}
########### 2. Manually set breakpoint at day8, then polynormial before and linear after.

# Define a custom variable for days before and after the breakpoint
species_totals$Days_poly <- pmin(species_totals$Days, 8)  # Polynomial part (before day 8)
species_totals$Days_linear <- pmax(species_totals$Days - 8, 0)  # Linear part (after day 8)


### poly 5: Fit the combined model
model_piecewise5 <- lm(log_Species_CFU ~ poly(Days_poly, 5)*trt*Species + Days_linear*trt*Species, data = species_totals)
summary(model_piecewise5)


### poly 3: Fit the combined model
model_piecewise3 <- lm(log_Species_CFU ~ poly(Days_poly, 3)*trt*Species + Days_linear*trt*Species, data = species_totals)
summary(model_piecewise3)


## visualization : Generate predictions for plotting
species_totals$Fitted <- predict(model_piecewise5)

# Plot the combined model
library(ggplot2)
ggplot(species_totals, aes(x = Days, y = log_Species_CFU)) +
  geom_point(alpha = 0.6, size = 1.2) +  # Observed data points
  geom_line(aes(y = Fitted), color = "blue", size = 1.2) +  # Fitted values
  labs(title = "Piecewise Model: Polynomial Before Day 8 and Linear After",
       x = "Days",
       y = "Log Species CFU") +
  geom_vline(xintercept = 8, linetype = "dashed", color = "red", size = 1) +  # Breakpoint
  theme_minimal()




###################### Compare:  ar1_poly_nested2,model_piecewise3, model_piecewise5
# compare poly 5 vs poly3
AIC(ar1_poly_nested2,segmented_model, model_piecewise3, model_piecewise5) # poly 5 is significantly better
anova(ar1_poly_nested2,segmented_model, model_piecewise3, model_piecewise5) # the difference between poly-3 and poly-5 is significant
anova(model_piecewise3, model_piecewise5)
```



##result 
```{r}
### poly 3: Fit the combined model
model_piecewise3 <- lm(log_Species_CFU ~ poly(Days_poly, 3)*trt*Species + Days_linear*trt*Species, data = species_totals)
summary(model_piecewise3)

library(lmerTest)
library(car)
anova_results <- Anova(model_piecewise3, type = "III")
print(anova_results)


```





```{r fig.height=8, fig.width=10}
check_model(model_piecewise3)
check_model(model_piecewise5)

```




## To test a random intercept and random slope model with a breakpoint at day 8 and a polynomial (degree 3) before the breakpoint and linear after,
```{r fig.height=8, fig.width=10}
breakpoint <- 8
species_totals$Days_poly <- ifelse(species_totals$Days <= breakpoint, species_totals$Days, 0)
species_totals$Days_linear <- ifelse(species_totals$Days > breakpoint, species_totals$Days - breakpoint, 0)

# Fit Random Intercept and Random Slope Model
library(lme4)
model_random_intercept <- lmer(
  log_Species_CFU ~ poly(Days_poly, 3) * trt * Species + Days_linear * trt * Species + (1 | ID),
  data = species_totals
)
summary(model_random_intercept)

#To account for subject-specific growth trends, include random slopes:
model_random_slope <- lmer(
  log_Species_CFU ~ poly(Days_poly, 3) * trt * Species + Days_linear * trt * Species + (1 + Days_poly + Days_linear | ID),
  data = species_totals
)
summary(model_random_slope)

# Compare models
anova(model_random_intercept, model_random_slope)
AIC(model_random_intercept, model_random_slope)
BIC(model_random_intercept, model_random_slope)
summary(model_random_slope)$varcor


# Diagnostic
library(ggplot2)
plot(residuals(model_random_slope), fitted(model_random_slope))
check_model(model_random_slope)

```





```{r}


```



