---
title: "Final_mydata_14counties"
author: "Jiseon Yang"
date: "2024-11-29"
output: word_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load data
```{r}


# Load libraries
library(sf)
library(spdep)
library(ggplot2)
library(dplyr)
library(viridis)
library(tmap)

# Load your dataset
final_data <- read.csv("C:/Users/jyang/OneDrive - Arizona State University/10 Classes_OneDrive/2024 8F_STP598_Spatiotemporal Analysis/Final Project/Final_data.csv")

# Preview the dataset structure
str(final_data)




```


### Step 2: Adjust Dataset for Spatial Analysis
```{r}

# Filter counties (excluding "Greenlee")
az_counties <- c(
  "Apache", "Cochise", "Coconino", "Gila", "Graham",
  "La Paz", "Maricopa", "Mohave", "Navajo", "Pima", 
  "Pinal", "Santa Cruz", "Yavapai", "Yuma"
)

# Verify the dataset includes only these counties
final_data <- final_data %>% filter(County %in% az_counties)

# Aggregate data (if needed, by Year and Month)
agg_data <- final_data %>%
  group_by(County, Year, Month) %>%
  summarize(
    TEMP_AVG = mean(TEMP_AVG, na.rm = TRUE),
    HUM_AVG = mean(HUM_AVG, na.rm = TRUE),
    WIND_AVG = mean(WIND_AVG, na.rm = TRUE),
    PM10 = mean(PM10, na.rm = TRUE),
    Cases = sum(Cases, na.rm = TRUE),
    .groups = "drop"
  )


```


### Step 3: Load and Prepare Shapefile
```{r}

# Load Arizona counties shapefile
shapefile_path <- "C:/Users/jyang/OneDrive - Arizona State University/10 Classes_OneDrive/2024 8F_STP598_Spatiotemporal Analysis/Final Project/tl_2024_us_county/az_counties.shp"
az_counties_sf <- st_read(shapefile_path)

# Filter shapefile to match dataset counties (excluding "Greenlee")
az_counties_sf <- az_counties_sf %>% filter(NAME %in% az_counties)

# Join aggregated data to shapefile
az_combined <- az_counties_sf %>%
  left_join(agg_data, by = c("NAME" = "County"))

# Check for mismatches in county names
setdiff(az_counties_sf$NAME, unique(agg_data$County))  # Counties in shapefile but not in agg_data
setdiff(unique(agg_data$County), az_counties_sf$NAME)  # Counties in agg_data but not in shapefile


```


### Step 4: Visualize Data
```{r}

# Plot PM10 levels
ggplot(data = az_combined) +
  geom_sf(aes(fill = PM10), color = "white") +
  scale_fill_viridis_c(option = "C", na.value = "grey50") +
  labs(title = "PM10 Levels by County",
       fill = "PM10 (µg/m³)") +
  theme_minimal()
```


### Step 5: Spatial Autocorrelation


```{r}

# Create neighbors list and weights matrix
nb <- poly2nb(az_counties_sf, queen = TRUE)
lw <- nb2listw(nb, style = "W")

# Moran's I for PM10
county_pm10 <- agg_data %>%
  group_by(County) %>%
  summarize(
    PM10_mean = mean(PM10, na.rm = TRUE),
    .groups = "drop"
  )

# Ensure order matches shapefile
county_pm10 <- county_pm10 %>%
  arrange(match(County, az_counties_sf$NAME))

# Run Moran's I
moran_test <- moran.test(county_pm10$PM10_mean, lw)
print(moran_test)


```


### Step 6: Temporal Analysis
```{r}

# Aggregate cases over time
time_series <- final_data %>%
  group_by(Year) %>%
  summarize(Cases = sum(Cases))

# Plot time series
ggplot(time_series, aes(x = Year, y = Cases)) +
  geom_line() +
  geom_point() +
  labs(title = "Disease Cases Over Time",
       x = "Year", y = "Cases")

# Aggregate cases over time
time_series <- final_data %>%
  group_by(Month) %>%
  summarize(Cases = sum(Cases))

# Plot time series
ggplot(time_series, aes(x = Month, y = Cases)) +
  geom_line() +
  geom_point() +
  labs(title = "Disease Cases Over Time",
       x = "Month", y = "Cases")

```


### Step 7: Regression Modeling
```{r}

# Linear model: PM10 and temperature predicting cases
lm_model <- lm(Cases ~ PM10 + TEMP_AVG + HUM_AVG + WIND_AVG, data = agg_data)
summary(lm_model)


```


### Step 8: Spatial Regression (Lag and Error Models)
```{r}
# Summarize PM10 and include other variables
county_pm10 <- agg_data %>%
  group_by(County) %>%
  summarize(
    PM10_mean = mean(PM10, na.rm = TRUE),
    TEMP_AVG = mean(TEMP_AVG, na.rm = TRUE),
    HUM_AVG = mean(HUM_AVG, na.rm = TRUE),
    WIND_AVG = mean(WIND_AVG, na.rm = TRUE),
    .groups = "drop"
  )

# Ensure county order matches the spatial weights matrix
county_pm10 <- county_pm10 %>%
  arrange(match(County, az_counties_sf$NAME))


library(spatialreg) 

# Spatial Lag Model
lag_model <- lagsarlm(PM10_mean ~ TEMP_AVG + HUM_AVG + WIND_AVG, data = county_pm10, listw = lw)

# Summarize the model
summary(lag_model)


# Spatial Error Model
error_model <- errorsarlm(PM10_mean ~ TEMP_AVG + HUM_AVG + WIND_AVG, data = county_pm10, listw = lw)
summary(error_model)

# Compare AIC values
AIC(lag_model, error_model)


```






### Step 9: Spatiotemporal Modeling with INLA


```{r}

library(INLA)
# Create county-level aggregated data
obs <- agg_data %>%
  group_by(County, Year) %>%
  summarize(
    PM10_mean = mean(PM10, na.rm = TRUE),
    TEMP_AVG = mean(TEMP_AVG, na.rm = TRUE),
    Cases = sum(Cases, na.rm = TRUE),
    .groups = "drop"
  )


# Create adjacency graph with correct order
nb <- poly2nb(az_counties_sf, queen = TRUE)
adj_matrix <- nb2mat(nb, style = "B", zero.policy = TRUE)
rownames(adj_matrix) <- az_counties_sf$NAME
graph <- inla.read.graph(adj_matrix)


# Convert County to a factor, ensuring levels match the adjacency matrix
obs$County <- factor(obs$County, levels = rownames(adj_matrix))


# Check for mismatches
setdiff(levels(obs$County), rownames(adj_matrix)) # Counties in data but not in adjacency matrix
setdiff(rownames(adj_matrix), levels(obs$County)) # Counties in adjacency matrix but not in data


# Ensure the graph's nodes match County levels
stopifnot(all(levels(obs$County) %in% rownames(adj_matrix)))

# Fit INLA model
# formula <- Cases ~ TEMP_AVG + PM10_mean +
#            f(County, model = "bym", graph = graph) +
#            f(Year, model = "rw1")

formula <- Cases ~ TEMP_AVG + PM10_mean +
           f(County, model = "iid") +  # Temporarily use iid for spatial effect
           f(Year, model = "rw1")     # Retain temporal random walk


fit_inla <- inla(
  formula,
  family = "poisson",
  data = obs,
  control.predictor = list(compute = TRUE)
)


# Summarize results
summary(fit_inla)


```


#### 




```{r}
print(adj_matrix)
```



```{r}
class(az_counties)  # Should include "sf"
colnames(az_counties)  # Should include "NAME" for county names

az_counties <- st_read("C:/Users/jyang/OneDrive - Arizona State University/10 Classes_OneDrive/2024 8F_STP598_Spatiotemporal Analysis/Final Project/tl_2024_us_county/az_counties.shp")

library(ggplot2)
library(sf)

# Ensure az_counties is an sf object with a column "NAME" for county names
ggplot(data = az_counties) +
  geom_sf(fill = "lightblue", color = "white") +  # Draw the map
  geom_sf_text(aes(label = NAME), size = 3, color = "black") +  # Add county names
  labs(title = "Arizona Counties",
       subtitle = "County names labeled on the map") +
  theme_minimal()

```





```{r}
# Load required libraries
library(sf)
library(dplyr)
library(ggplot2)
library(viridis)
library(tmap)
library(INLA)
library(fields)
library(spdep)

# Load your dataset
final_data <- read.csv("C:/Users/jyang/OneDrive - Arizona State University/10 Classes_OneDrive/2024 8F_STP598_Spatiotemporal Analysis/Final Project/Final_data.csv")

# Preview the dataset structure
str(final_data)

# Define the list of Arizona counties
az_counties <- c(
  "Apache", "Cochise", "Coconino", "Gila", "Graham",
  "La Paz", "Maricopa", "Mohave", "Navajo", "Pima",
  "Pinal", "Santa Cruz", "Yavapai", "Yuma"
)

# Filter the dataset to include only these counties
final_data <- final_data %>% filter(County %in% az_counties)

# Aggregate the data by Year and Month
agg_data <- final_data %>%
  group_by(County, Year, Month) %>%
  summarize(
    TEMP_AVG = mean(TEMP_AVG, na.rm = TRUE),
    HUM_AVG = mean(HUM_AVG, na.rm = TRUE),
    WIND_AVG = mean(WIND_AVG, na.rm = TRUE),
    PM10 = mean(PM10, na.rm = TRUE),
    Cases = sum(Cases, na.rm = TRUE),
    .groups = "drop"
  )

# Load Arizona counties shapefile
shapefile_path <- "C:/Users/jyang/OneDrive - Arizona State University/10 Classes_OneDrive/2024 8F_STP598_Spatiotemporal Analysis/Final Project/tl_2024_us_county/az_counties.shp"
az_counties_sf <- st_read(shapefile_path)

# Filter shapefile to match the dataset counties
az_counties_sf <- az_counties_sf %>% filter(NAME %in% az_counties)

# Join aggregated data with the shapefile
az_combined <- az_counties_sf %>%
  left_join(agg_data, by = c("NAME" = "County"))

# Ensure there are no mismatches in the join
setdiff(az_counties_sf$NAME, unique(agg_data$County))  # Should return character(0)
setdiff(unique(agg_data$County), az_counties_sf$NAME)  # Should return character(0)

# Create neighbors list and spatial weights matrix
nb <- poly2nb(az_counties_sf, queen = TRUE)
lw <- nb2listw(nb, style = "W")

# Plot PM10 levels across counties
ggplot(data = az_combined) +
  geom_sf(aes(fill = PM10), color = "white") +
  scale_fill_viridis_c(option = "C", na.value = "grey50") +
  labs(title = "PM10 Levels by County",
       fill = "PM10 (µg/m³)") +
  theme_minimal()
```


```{r}
# Temporal trend plot for disease cases
library(gridExtra)

# Plot for Yearly Cases
time_series <- final_data %>%
  group_by(Year) %>%
  summarize(Cases = sum(Cases))


plot_year <- ggplot(time_series, aes(x = Year, y = Cases)) +
  geom_line() +
  geom_point() +
  labs(title = "Disease Cases Over Time",
       x = "Yearh", y = "Cases")

# Plot for Monthly Cases
time_series <- final_data %>%
  group_by(Month) %>%
  summarize(Cases = sum(Cases))

plot_month <- ggplot(time_series, aes(x = Month, y = Cases)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(1, 12, by = 1)) + # Ensures every month is displayed
  labs(title = "Disease Cases Over Time",
       x = "Month", y = "Cases")

grid.arrange(plot_year, plot_month, ncol = 2)

```



```{r}
# Prepare for Spatiotemporal Modeling with INLA
obs <- agg_data %>%
  group_by(County, Year) %>%
  summarize(
    PM10_mean = mean(PM10, na.rm = TRUE),
    TEMP_AVG = mean(TEMP_AVG, na.rm = TRUE),
    Cases = sum(Cases, na.rm = TRUE),
    .groups = "drop"
  )

# Convert County to a factor and ensure levels match adjacency matrix
obs$County <- factor(obs$County, levels = az_counties_sf$NAME)

# Ensure adjacency matrix matches the data
adj_matrix <- nb2mat(nb, style = "B", zero.policy = TRUE)
rownames(adj_matrix) <- az_counties_sf$NAME
graph <- inla.read.graph(adj_matrix)

# Fit INLA spatiotemporal model
formula <- Cases ~ TEMP_AVG + PM10_mean +
           f(County, model = "iid") +  # Temporarily use iid for spatial effect
           f(Year, model = "rw1")     # Retain temporal random walk

fit_inla <- inla(
  formula,
  family = "poisson",
  data = obs,
  control.predictor = list(compute = TRUE)
)

# Summarize INLA results
summary(fit_inla)

```





```{r}

# Add predicted cases to the dataset for visualization
obs$Cases_pred <- fit_inla$summary.fitted.values$mean

# Join predicted cases back to the spatial dataset
az_combined <- az_combined %>%
  left_join(obs %>% select(County, Cases_pred), by = c("NAME" = "County"))

# Plot predicted cases
ggplot(data = az_combined) +
  geom_sf(aes(fill = Cases_pred), color = "white") +
  scale_fill_viridis_c(option = "C", na.value = "grey50") +
  labs(title = "Predicted Cases by County",
       fill = "Predicted Cases") +
  theme_minimal()

```

################################################
#Impact of Seasonality on Research Modeling

1. Time Series Models :ARIMA with Seasonality
```{r}
# install.packages("forecast")
library(forecast)

maricopa_pm10 <- ts(final_data$PM10[final_data$County == "Maricopa"], frequency = 12)
sarima_model <- auto.arima(maricopa_pm10, seasonal = TRUE)
summary(sarima_model)


library(astsa)



# Plot and decompose
plot(maricopa_pm10, main = "PM10 Levels in Maricopa")
decompose(maricopa_pm10)

```


2. Spatiotemporal Models (e.g., INLA)
```{r}
obs <- agg_data %>%
  group_by(County, Year, Month) %>%  # Include Month in the grouping
  summarize(
    PM10_mean = mean(PM10, na.rm = TRUE),
    TEMP_AVG = mean(TEMP_AVG, na.rm = TRUE),
    Cases = sum(Cases, na.rm = TRUE),
    .groups = "drop"
  )

obs$Month <- as.numeric(obs$Month)  # Ensure Month is numeric
obs$Month <- factor(obs$Month, levels = 1:12)  # Convert to factor if required

formula <- Cases ~ TEMP_AVG + PM10_mean +
           f(County, model = "iid") +  # Spatial effect
           f(Year, model = "rw1") +   # Long-term trend
           f(Month, model = "rw1", cyclic = TRUE)  # Seasonal effect

fit_inla <- inla(
  formula,
  family = "poisson",
  data = obs,
  control.predictor = list(compute = TRUE)
)

summary(fit_inla)


```


Visualize Predicted Cases
```{r}
obs$Cases_pred <- fit_inla$summary.fitted.values$mean

# Join predictions back to spatial data
az_combined <- az_counties_sf %>%
  left_join(obs, by = c("NAME" = "County"))

# Plot predicted cases
ggplot(data = az_combined) +
  geom_sf(aes(fill = Cases_pred), color = "white") +
  scale_fill_viridis_c(option = "C", na.value = "grey50") +
  labs(title = "Predicted Cases with Seasonal Effect",
       fill = "Predicted Cases") +
  theme_minimal()

```



1. ARIMA Model for Temporal Patterns in PM10
```{r}

# Subset PM10 time series for Maricopa County
maricopa_pm10 <- ts(obs$PM10_mean[obs$County == "Maricopa"], frequency = 12, start = c(2000, 1))

# Fit a SARIMA model
sarima_model <- astsa::sarima(maricopa_pm10, p = 1, d = 1, q = 1, P = 1, D = 1, Q = 1, S = 12)

# View results
print(sarima_model)


##########

library(forecast)


# Fit SARIMA model
sarima_model <- auto.arima(maricopa_pm10, seasonal = TRUE)
summary(sarima_model)

# Decompose and visualize
decomposed <- decompose(maricopa_pm10)
plot(decomposed)


```



2. Spatiotemporal Analysis with the SpatioTemporal Package
```{r}
# install.packages("SpatioTemporal")
library(SpatioTemporal)

str(obs)

library(dplyr)  # For dplyr functions, including `%>%`
library(tidyverse)  # For the entire tidyverse, which includes dplyr

# Check missing data percentage
sapply(obs, function(x) sum(is.na(x) | is.nan(x)) / length(x) * 100)



# # Option 1: Remove rows with NaN values
# obs <- obs %>%
#   filter(!is.nan(PM10_mean) & !is.nan(TEMP_AVG))

# Option 2: Impute missing values (e.g., with mean or interpolation)
library(zoo)
obs <- obs %>%
  mutate(
    PM10_mean = na.approx(PM10_mean, na.rm = FALSE),
    TEMP_AVG = na.approx(TEMP_AVG, na.rm = FALSE)
  )
obs <- obs %>%
  mutate(Date = as.Date(paste(Year, Month, "01", sep = "-")))
obs$PM10_mean <- as.numeric(obs$PM10_mean)
obs$TEMP_AVG <- as.numeric(obs$TEMP_AVG)

# If not already created, create combined_data
combined_data <- obs %>%
  group_by(County) %>%
  summarize(
    TEMP_AVG = mean(TEMP_AVG, na.rm = TRUE),
    PM10_mean = mean(PM10_mean, na.rm = TRUE),
    Cases = sum(Cases, na.rm = TRUE),
    .groups = "drop"
  )

# If combined_data does not yet have spatial geometry, ensure az_counties_sf is available
library(sf)
combined_data <- az_counties_sf %>%
  left_join(combined_data, by = c("NAME" = "County"))

# Convert to Spatial object
library(sp)
combined_data_sp <- as(combined_data, "Spatial")

# Ensure the column `ID` exists in `combined_data_sp`
combined_data_sp$ID <- combined_data_sp$NAME  # Copy county names as IDs if necessary
setdiff(obs$County, combined_data_sp$ID)



# Check mismatches
setdiff(obs$County, combined_data_sp$ID)  # Or whatever column identifies locations in `combined_data_sp`

# Ensure consistent IDs
obs$County <- factor(obs$County, levels = unique(combined_data_sp$ID))

str(obs)

obs <- obs %>%
  rename(
    obs = PM10_mean,   # Observed values (PM10_mean for this example)
    date = Date,       # Ensure Date column is named `date`
    ID = County        # Ensure County column is named `ID`
  )

combined_data_sp$ID <- combined_data_sp$NAME  # Replace NAME with the spatial identifier column

covars <- obs %>%
  group_by(ID) %>%
  summarize(
    TEMP_AVG = mean(TEMP_AVG, na.rm = TRUE),
    Cases = sum(Cases, na.rm = TRUE),
    .groups = "drop"
  )


data_st <- createSTdata(
  obs = obs,
  spLocations = combined_data_sp,
  Time = unique(obs$date),
  Variables = c("TEMP_AVG", "Cases"),
  covars = covars
)



summary(data_st)


```



```{r}


```



```{r}


```



```{r}


```




