---
title: "Micro-5_TD50_JY"
author: "Jiseon Yang"
date: "2024-11-19"
output: word_document
---


```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(car)  # For ANOVA assumption testing
library(multcomp)  # For Tukey's HSD

# Load dataset (replace 'td50_data.csv' with your file path)
data <- read.csv("C:/Users/jyang/ASU Dropbox/Jiseon Yang/2 PROJJECT/2014_NASA Micro-5/Micro-17_TD50.csv")

# Check data structure
str(data)


```




```{r}
# Convert categorical variables to factors
data$ID <- as.factor(data$Samplenumber)
data$Strain <- as.factor(data$Strain)
data$Environment <- as.factor(data$GravityCondition)
data$Media <- as.factor(data$Media)

# Summary statistics
summary(data)
summary(data$TD50)

# Remove rows with missing data
data_clean <- na.omit(data)
```

# Data overview

```{r }

# Boxplots to visualize TD50 distribution
# ggplot(data_clean, aes(x=Strain, y=TD50, fill=Environment)) +
#   geom_boxplot() +
#   facet_wrap(~ Media) +
#   theme_minimal() +
#   labs(title="TD50 by Strain, Environment, and Media",
#        x="Strain", y="TD50")

# Load necessary library
library(gridExtra)
library(ggplot2)

# Define a function to plot for a specific media type without a legend
plot_media <- function(media_name, legend = FALSE) {
  p <- ggplot(data_clean[data_clean$Media == media_name, ], aes(x = Strain, y = TD50, fill = Environment)) +
    geom_boxplot() +
    theme_minimal() +
    labs(title = paste("TD50:", media_name),
         x = NULL, y = "TD50") +
    scale_x_discrete(limits = c("OP50", "Salmonella")) +
    scale_y_continuous(limits = c(4, 15), breaks = seq(4, 16, by = 2)) +
    theme(
      plot.title = element_text(size = 12, face = "bold"),#font size and style
      axis.title.x = element_text(size = 9),   # X-axis title font size
      axis.title.y = element_text(size = 9),   # Y-axis title font size
      axis.text.x = element_text(size = 9),    # X-axis tick labels font size
      axis.text.y = element_text(size = 9)     # Y-axis tick labels font size
    )
  if (!legend) {
    p <- p + theme(legend.position = "none")  # Remove legend
  }
  return(p)
}

# Generate individual plots
plot_LB <- plot_media("LB", legend = FALSE)
plot_LBM9 <- plot_media("LBM9", legend = FALSE)
plot_LBPiPEG <- plot_media("LBPiPEG", legend = FALSE)
plot_LBPO4 <- plot_media("LBPO4", legend = FALSE)  # Include legend for this plot

# Extract the legend from one plot
get_legend <- function(plot) {
  g <- ggplotGrob(plot)
  legend <- g$grobs[which(sapply(g$grobs, function(x) x$name) == "guide-box")][[1]]
  return(legend)
}

# Arrange plots in a 2x2 grid with a shared legend
grid.arrange(
  arrangeGrob(plot_LB, plot_LBM9, plot_LBPiPEG, plot_LBPO4, nrow = 2, ncol = 2),
  legend,
  nrow = 2,
  heights = c(3, 1)  # Adjust height ratio between plots and legend
)
```

# Flight vs Ground Static
```{r }
 # Define a function to plot based on Environment and Strain
plot_environment_strain <- function(environment, strain) {
  ggplot(data_clean[data_clean$Environment == environment & data_clean$Strain == strain, ],
         aes(x = Media, y = TD50, fill = Media)) +
    geom_boxplot() +
    theme_minimal() +
    labs(title = paste(environment, "-", strain),
         x = "Media", y = "TD50") +
    scale_y_continuous(limits = c(4, 13), breaks = seq(4, 13, by = 2)) +
    theme(
      plot.title = element_text(size = 12, face = "bold"),
      axis.title.x = element_text(size = 9),
      axis.title.y = element_text(size = 9),
      axis.text.x = element_text(size = 8, angle = 45, hjust = 1),  # Rotate x-axis labels
      axis.text.y = element_text(size = 9),
      legend.position = "none"  # Remove individual legends
    )
}

# Generate the plots for each combination of Environment and Strain
plot_flight_op50 <- plot_environment_strain("Flight", "OP50")
plot_flight_salmonella <- plot_environment_strain("Flight", "Salmonella")
plot_ground_static_op50 <- plot_environment_strain("Ground Static", "OP50")
plot_ground_static_salmonella <- plot_environment_strain("Ground Static", "Salmonella")


# Combine the plots into a 2x2 grid
library(gridExtra)

grid.arrange(
  arrangeGrob(
    plot_flight_op50, plot_flight_salmonella,
    plot_ground_static_op50, plot_ground_static_salmonella,
    nrow = 2, ncol = 2
  ),
  nrow = 1
)
```




```{r}

# Interaction plots
interaction.plot(data_clean$Strain, data_clean$Environment, data_clean$TD50, col=c("red", "blue"))
interaction.plot(data_clean$Media, data_clean$Environment, data_clean$TD50, col=c("green", "purple"))


```

# ANOVA
\n Conclusion: Strain, Environment, and the interaction (Strain* Environment) significantly affect TD50.
\n violate the assumption of residual is not normal
```{r}
# ANOVA model
anova_model <- aov(TD50 ~ Strain * Environment * Media, data=data_clean)

# ANOVA table
summary(anova_model)


# Post-hoc test
#tukey_result <- TukeyHSD(anova_model)

# Display Tukey's HSD results
#print(tukey_result)


# Save results to a file
#write.csv(tukey_result$`Strain:Environment:Media`, "tukey_results.csv")


# Homogeneity of variances
# H0: variances are equal across groups
leveneTest(TD50 ~ Strain * Environment * Media, data=data_clean)
cat("\n * There is no evidence of significant heteroscedasticity:variances are homogeneous")

# Check assumptions of ANOVA
# Residuals normality: H0: residual~ Normal
shapiro.test(residuals(anova_model))
cat("* The residuals are not normally distributed.","\n" )

# Transform and check normality again
data_clean$TD50_transformed <- log(data_clean$TD50)
anova_model_transformed <- aov(TD50_transformed ~ Strain * Environment * Media, data = data_clean)
shapiro.test(residuals(anova_model_transformed))
cat("* The residuals are still not normally distributed.","\n" )

# Conclusion:
cat("\n\n Conclusion:")
cat("\n 1. Strain, Environment, and the interaction (Strain* Environment) significantly affect TD50.")
cat("\n 2. Media and its interactions do not significantly influence TD50.")
cat("\n 3. Variance homogeneity is met, and the ANOVA results are reliable. Yet, residual is not normal. Try nonparametric.")
cat("\n 4. Insight: Focus on Strain and Environment for their significant effects on TD50.")
```




# Test Nonparametric Methods: Kruskal-Wallis rank sum test
\n H0: The medians of TD50 are the same across all combinations of Strain, Environment, and Media.
\n strong evidence suggesting that at least one combination of factors significantly affects TD50.
```{r}
# test Median
kruskal.test(TD50 ~ interaction(Strain, Environment, Media), data = data_clean)
cat("\n * we reject the null hypothesis. There is strong evidence that at least one combination of factors significantly affects TD50." )


library(ggplot2)
ggplot(data_clean, aes(x = interaction(Strain, Environment, Media), y = TD50, fill = Environment)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "TD50 by Strain, Environment, and Media",
    x = "Factor Combinations",
    y = "TD50"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

```




# generalized linear mode:   
\n Parametric modeling, but it can be robust to non-normal residuals
## GLM_Gaussian vs GLM_Gamma: 
\n Better fit with gamma distribution ~
\n non-negative, continuous, may skewed like TD50, variance linked to mean (larger values often show greater variability)
```{r}
glm_gaussian <- glm(TD50 ~ Strain * Environment * Media, family = gaussian(link = "log"), data = data_clean)
summary(glm_gaussian)
anova(glm_gaussian)


cat("\n\n Conclusion:")
cat("\n 1. The results indicate that the type of Strain (OP50 vs. Salmonella) and the Environment (Flight vs. Ground Rotating or Ground Static) are the primary factors influencing TD50.")
cat("\n 2. Media and interactions among factors do not play a significant role in determining TD50 under the conditions tested.")

# gamma: non-negative, continuous, may skewed like TD50, variance linked to mean (larger values often show greater variability)
glm_Gamma <- glm(TD50 ~ Strain * Environment * Media, family = Gamma(link = "log"), data = data_clean)
summary(glm_Gamma)
anova(glm_Gamma)

cat("\n\n Conclusion:")
cat("\n Strain and Environment are the primary factors influencing TD50.")
cat("\n Environment*Media: environmental effects on TD50 depend on the media type.  ")
cat("\n   ")

# Model Diagnostics for Gamma GLM
par(mfrow = c(2, 2))
plot(glm_Gamma)
data_clean[c(64, 70, 71), "Samplenumber"]

# Compare AIC
AIC(glm_gaussian, glm_Gamma)
cat("\n glm_Gamma is better model  ")

```
\n Conclusion:
\n 1. The results indicate that the type of Strain (OP50 vs. Salmonella) and the Environment (Flight vs. Ground Rotating or Ground Static) are the primary factors influencing TD50.
\n 2. Media and interactions among factors do not play a significant role in determining TD50 under the conditions tested.




# Perform Aligned Rank Transform (ART) ANOVA,

#### The Aligned Rank Transform (ART) ANOVA is a nonparametric statistical method used to analyze factorial designs, especially when data violate the assumptions of traditional ANOVA. It extends the utility of nonparametric methods, like rank-based tests, to handle interactions and complex designs.
\n•Standard ANOVA assumes normality and homogeneity of variance; ART ANOVA does not.
\n•While ANOVA analyzes means, ART ANOVA analyzes the ranked equivalent of aligned data.
```{r}
#install.packages("remotes")  
remotes::install_github("mjskay/ARTool")
library(ARTool)
art_model <- art(TD50 ~ Strain * Environment * Media, data = data_clean)

anova(art_model)

```


\n Interpretation
\n 1.	Main Effects:
\n o	Strain, Environment, and Media independently influence TD50.
\n o	Environment has the strongest impact, followed by Strain.
\n 2.	Two-Way Interactions:
\n o	The effect of Strain on TD50 depends on the Environment.
\n o	The effect of Strain depends on the Media.
\n o	The effect of Environment depends on the Media.
\n 3.	Three-Way Interaction:
\n o	The combined effect of Strain, Environment, and Media is not significant, suggesting no additional complexity beyond the two-way interactions.








# Remove outliers
```{r}
data_clean[c(64, 70, 71), ]

# Remove outliers
data_no_outliers <- data_clean[-c(64, 70, 71), ]





########### re-plot
# Define a function to plot for a specific media type without a legend
plot_media <- function(media_name, legend = FALSE) {
  p <- ggplot(data_no_outliers[data_no_outliers$Media == media_name, ], aes(x = Strain, y = TD50, fill = Environment)) +
    geom_boxplot() +
    theme_minimal() +
    labs(title = paste("TD50:", media_name),
         x = NULL, y = "TD50") +
    scale_x_discrete(limits = c("OP50", "Salmonella")) +
    scale_y_continuous(limits = c(4, 15), breaks = seq(4, 16, by = 2)) +
    theme(
      plot.title = element_text(size = 12, face = "bold"),#font size and style
      axis.title.x = element_text(size = 9),   # X-axis title font size
      axis.title.y = element_text(size = 9),   # Y-axis title font size
      axis.text.x = element_text(size = 9),    # X-axis tick labels font size
      axis.text.y = element_text(size = 9)     # Y-axis tick labels font size
    )
  if (!legend) {
    p <- p + theme(legend.position = "none")  # Remove legend
  }
  return(p)
}

# Generate individual plots
plot_LB <- plot_media("LB", legend = FALSE)
plot_LBM9 <- plot_media("LBM9", legend = FALSE)
plot_LBPiPEG <- plot_media("LBPiPEG", legend = FALSE)
plot_LBPO4 <- plot_media("LBPO4", legend = FALSE)  # Include legend for this plot

# Extract the legend from one plot
get_legend <- function(plot) {
  g <- ggplotGrob(plot)
  legend <- g$grobs[which(sapply(g$grobs, function(x) x$name) == "guide-box")][[1]]
  return(legend)
}

# Arrange plots in a 2x2 grid with a shared legend
grid.arrange(
  arrangeGrob(plot_LB, plot_LBM9, plot_LBPiPEG, plot_LBPO4, nrow = 2, ncol = 2),
  legend,
  nrow = 2,
  heights = c(3, 1)  # Adjust height ratio between plots and legend
)




################# re-plot

 # Define a function to plot based on Environment and Strain
plot_environment_strain <- function(environment, strain) {
  ggplot(data_no_outliers[data_no_outliers$Environment == environment & data_no_outliers$Strain == strain, ],
         aes(x = Media, y = TD50, fill = Media)) +
    geom_boxplot() +
    theme_minimal() +
    labs(title = paste(environment, "-", strain),
         x = "Media", y = "TD50") +
    scale_y_continuous(limits = c(4, 13), breaks = seq(4, 13, by = 2)) +
    theme(
      plot.title = element_text(size = 12, face = "bold"),
      axis.title.x = element_text(size = 9),
      axis.title.y = element_text(size = 9),
      axis.text.x = element_text(size = 8, angle = 45, hjust = 1),  # Rotate x-axis labels
      axis.text.y = element_text(size = 9),
      legend.position = "none"  # Remove individual legends
    )
}

# Generate the plots for each combination of Environment and Strain
plot_flight_op50 <- plot_environment_strain("Flight", "OP50")
plot_flight_salmonella <- plot_environment_strain("Flight", "Salmonella")
plot_ground_static_op50 <- plot_environment_strain("Ground Static", "OP50")
plot_ground_static_salmonella <- plot_environment_strain("Ground Static", "Salmonella")


# Combine the plots into a 2x2 grid
library(gridExtra)

grid.arrange(
  arrangeGrob(
    plot_flight_op50, plot_flight_salmonella,
    plot_ground_static_op50, plot_ground_static_salmonella,
    nrow = 2, ncol = 2
  ),
  nrow = 1
)


####################### re-plot
library(ggplot2)
ggplot(data_no_outliers, aes(x = interaction(Strain, Environment, Media), y = TD50, fill = Environment)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "TD50 by Strain, Environment, and Media",
    x = "Factor Combinations",
    y = "TD50"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```



```{r}
glm_no_outliers <- glm(TD50 ~ Strain * Environment * Media, family = Gamma(link = "log"), data = data_no_outliers)

summary(glm_no_outliers)

anova(glm_no_outliers)

par(mfrow = c(2, 2))
plot(glm_no_outliers)

# Compare AIC
AIC(glm_no_outliers, glm_Gamma)
cat("\n glm model: Better without outliers")
```



```{r}
art_model_Outliers <- art(TD50 ~ Strain * Environment * Media, data = data_no_outliers)

anova(art_model_Outliers)

```






# Remove selected data : 
## Indeed, removing the selected data (odd data) better satisfy assumption although the overall pattern may not agree accross all conditions. 
```{r}
data_clean[c(60, 63, 66, 69), ]

# Remove outliers
data_remove_select <- data_clean[-c(60, 63, 66, 69), ]



########### re-plot
# Define a function to plot for a specific media type without a legend
plot_media <- function(media_name, legend = FALSE) {
  p <- ggplot(data_remove_select[data_remove_select$Media == media_name, ], aes(x = Strain, y = TD50, fill = Environment)) +
    geom_boxplot() +
    theme_minimal() +
    labs(title = paste("TD50:", media_name),
         x = NULL, y = "TD50") +
    scale_x_discrete(limits = c("OP50", "Salmonella")) +
    scale_y_continuous(limits = c(4, 15), breaks = seq(4, 16, by = 2)) +
    theme(
      plot.title = element_text(size = 12, face = "bold"),#font size and style
      axis.title.x = element_text(size = 9),   # X-axis title font size
      axis.title.y = element_text(size = 9),   # Y-axis title font size
      axis.text.x = element_text(size = 9),    # X-axis tick labels font size
      axis.text.y = element_text(size = 9)     # Y-axis tick labels font size
    )
  if (!legend) {
    p <- p + theme(legend.position = "none")  # Remove legend
  }
  return(p)
}

# Generate individual plots
plot_LB <- plot_media("LB", legend = FALSE)
plot_LBM9 <- plot_media("LBM9", legend = FALSE)
plot_LBPiPEG <- plot_media("LBPiPEG", legend = FALSE)
plot_LBPO4 <- plot_media("LBPO4", legend = FALSE)  # Include legend for this plot

# Extract the legend from one plot
get_legend <- function(plot) {
  g <- ggplotGrob(plot)
  legend <- g$grobs[which(sapply(g$grobs, function(x) x$name) == "guide-box")][[1]]
  return(legend)
}

# Arrange plots in a 2x2 grid with a shared legend
grid.arrange(
  arrangeGrob(plot_LB, plot_LBM9, plot_LBPiPEG, plot_LBPO4, nrow = 2, ncol = 2),
  legend,
  nrow = 2,
  heights = c(3, 1)  # Adjust height ratio between plots and legend
)




################# re-plot

 # Define a function to plot based on Environment and Strain
plot_environment_strain <- function(environment, strain) {
  ggplot(data_remove_select[data_remove_select$Environment == environment & data_remove_select$Strain == strain, ],
         aes(x = Media, y = TD50, fill = Media)) +
    geom_boxplot() +
    theme_minimal() +
    labs(title = paste(environment, "-", strain),
         x = "Media", y = "TD50") +
    scale_y_continuous(limits = c(4, 13), breaks = seq(4, 13, by = 2)) +
    theme(
      plot.title = element_text(size = 12, face = "bold"),
      axis.title.x = element_text(size = 9),
      axis.title.y = element_text(size = 9),
      axis.text.x = element_text(size = 8, angle = 45, hjust = 1),  # Rotate x-axis labels
      axis.text.y = element_text(size = 9),
      legend.position = "none"  # Remove individual legends
    )
}

# Generate the plots for each combination of Environment and Strain
plot_flight_op50 <- plot_environment_strain("Flight", "OP50")
plot_flight_salmonella <- plot_environment_strain("Flight", "Salmonella")
plot_ground_static_op50 <- plot_environment_strain("Ground Static", "OP50")
plot_ground_static_salmonella <- plot_environment_strain("Ground Static", "Salmonella")


# Combine the plots into a 2x2 grid
library(gridExtra)

grid.arrange(
  arrangeGrob(
    plot_flight_op50, plot_flight_salmonella,
    plot_ground_static_op50, plot_ground_static_salmonella,
    nrow = 2, ncol = 2
  ),
  nrow = 1
)


####################### re-plot
library(ggplot2)
ggplot(data_remove_select, aes(x = interaction(Strain, Environment, Media), y = TD50, fill = Environment)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "TD50 by Strain, Environment, and Media",
    x = "Factor Combinations",
    y = "TD50"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```


## Diagnostic with data_remove_select
\n Overall, satisfy assumption.
\n Could remove further data c(71, 40, 12)  but not necessary
```{r fig.height= 8}
# Diagnostics
par(mfrow = c(2, 2))
plot(glm_remove_select)

# non-linearity and heteroscedasticity (non-constant variance)
# residuals are approximately Gamma-distributed, with slight deviations in the tails. -->  acceptable
# homoscedasticity (constant variance): fairly consistent across fitted values
# possible leverage point: 71, close to the Cook's distance threshold, may have some influence on the model.
# 48 and 12, could have moderate leverage but within threshold 

data_clean[c(71, 40, 12), ]

```




## Fitting glm with data_remove_select
```{r }

glm_remove_select <- glm(TD50 ~ Strain * Environment * Media, family = Gamma(link = "log"), data = data_remove_select)

summary(glm_remove_select)

anova(glm_remove_select)


# Compare AIC
AIC(glm_no_outliers, glm_Gamma, glm_remove_select)
```


## Fitting art with data_remove_select

```{r}
art_remove_select <- art(TD50 ~ Strain * Environment * Media, data = data_remove_select)

anova(art_remove_select)

```
\n Conclusion:

\n •	Both models identify the same significant main effects and interactions, including Strain, Environment, Media, and their two-way interactions. 
\n •	The three-way interaction is borderline significant in the ART model and significant in the Gamma GLM.




# Ground Rotating vs Ground Static
```{r }
 # Define a function to plot based on Environment and Strain
plot_environment_strain <- function(environment, strain) {
  ggplot(data_clean[data_clean$Environment == environment & data_clean$Strain == strain, ],
         aes(x = Media, y = TD50, fill = Media)) +
    geom_boxplot() +
    theme_minimal() +
    labs(title = paste(environment, "-", strain),
         x = "Media", y = "TD50") +
    scale_y_continuous(limits = c(4, 15), breaks = seq(4, 15, by = 2)) +
    theme(
      plot.title = element_text(size = 12, face = "bold"),
      axis.title.x = element_text(size = 9),
      axis.title.y = element_text(size = 9),
      axis.text.x = element_text(size = 8, angle = 45, hjust = 1),  # Rotate x-axis labels
      axis.text.y = element_text(size = 9),
      legend.position = "none"  # Remove individual legends
    )
}

# Generate the plots for each combination of Environment and Strain
plot_flight_op50 <- plot_environment_strain("Ground Rotating", "OP50")
plot_flight_salmonella <- plot_environment_strain("Ground Rotating", "Salmonella")
plot_ground_static_op50 <- plot_environment_strain("Ground Static", "OP50")
plot_ground_static_salmonella <- plot_environment_strain("Ground Static", "Salmonella")


# Combine the plots into a 2x2 grid
library(gridExtra)

grid.arrange(
  arrangeGrob(
    plot_flight_op50, plot_flight_salmonella,
    plot_ground_static_op50, plot_ground_static_salmonella,
    nrow = 2, ncol = 2
  ),
  nrow = 1
)

################## data_no_outliers
 # Define a function to plot based on Environment and Strain
plot_environment_strain <- function(environment, strain) {
  ggplot(data_no_outliers[data_no_outliers$Environment == environment & data_no_outliers$Strain == strain, ],
         aes(x = Media, y = TD50, fill = Media)) +
    geom_boxplot() +
    theme_minimal() +
    labs(title = paste(environment, "-", strain),
         x = "Media", y = "TD50") +
    scale_y_continuous(limits = c(4, 15), breaks = seq(4, 15, by = 2)) +
    theme(
      plot.title = element_text(size = 12, face = "bold"),
      axis.title.x = element_text(size = 9),
      axis.title.y = element_text(size = 9),
      axis.text.x = element_text(size = 8, angle = 45, hjust = 1),  # Rotate x-axis labels
      axis.text.y = element_text(size = 9),
      legend.position = "none"  # Remove individual legends
    )
}

# Generate the plots for each combination of Environment and Strain
plot_flight_op50 <- plot_environment_strain("Ground Rotating", "OP50")
plot_flight_salmonella <- plot_environment_strain("Ground Rotating", "Salmonella")
plot_ground_static_op50 <- plot_environment_strain("Ground Static", "OP50")
plot_ground_static_salmonella <- plot_environment_strain("Ground Static", "Salmonella")


# Combine the plots into a 2x2 grid
library(gridExtra)

grid.arrange(
  arrangeGrob(
    plot_flight_op50, plot_flight_salmonella,
    plot_ground_static_op50, plot_ground_static_salmonella,
    nrow = 2, ncol = 2
  ),
  nrow = 1
)


################## data_remove_select
 # Define a function to plot based on Environment and Strain
plot_environment_strain <- function(environment, strain) {
  ggplot(data_remove_select[data_remove_select$Environment == environment & data_remove_select$Strain == strain, ],
         aes(x = Media, y = TD50, fill = Media)) +
    geom_boxplot() +
    theme_minimal() +
    labs(title = paste(environment, "-", strain),
         x = "Media", y = "TD50") +
    scale_y_continuous(limits = c(4, 15), breaks = seq(4, 15, by = 2)) +
    theme(
      plot.title = element_text(size = 12, face = "bold"),
      axis.title.x = element_text(size = 9),
      axis.title.y = element_text(size = 9),
      axis.text.x = element_text(size = 8, angle = 45, hjust = 1),  # Rotate x-axis labels
      axis.text.y = element_text(size = 9),
      legend.position = "none"  # Remove individual legends
    )
}

# Generate the plots for each combination of Environment and Strain
plot_flight_op50 <- plot_environment_strain("Ground Rotating", "OP50")
plot_flight_salmonella <- plot_environment_strain("Ground Rotating", "Salmonella")
plot_ground_static_op50 <- plot_environment_strain("Ground Static", "OP50")
plot_ground_static_salmonella <- plot_environment_strain("Ground Static", "Salmonella")


# Combine the plots into a 2x2 grid
library(gridExtra)

grid.arrange(
  arrangeGrob(
    plot_flight_op50, plot_flight_salmonella,
    plot_ground_static_op50, plot_ground_static_salmonella,
    nrow = 2, ncol = 2
  ),
  nrow = 1
)
```



# Exclude PiPEG

```{r}
# Remove rows where Media == "LBPiPEG"
data_no_LBPiPEG <- data_remove_select[data_remove_select$Media != "LBPiPEG", ]


# Gamma GLM
glm_no_LBPiPEG <- glm(
  TD50 ~ Strain * Environment * Media,
  family = Gamma(link = "log"),
  data = data_no_LBPiPEG
)
summary(glm_no_LBPiPEG)

# Diagnostic plots
par(mfrow = c(2, 2))
plot(glm_no_LBPiPEG)  

#Type III ANOVA
library(car)
type_iii_anova_no_LBPiPEG <- Anova(glm_no_LBPiPEG, type = "III")
print(type_iii_anova_no_LBPiPEG)

#Type I anova
anova(glm_no_LBPiPEG)

library(ggplot2)
ggplot(data_no_LBPiPEG, aes(x = Media, y = TD50, fill = Environment)) +
  geom_boxplot() +
  facet_wrap(~ Strain) +
  theme_minimal() +
  labs(title = "TD50 by Strain, Environment, and Media (Excluding LBPiPEG)")
```






# Exclude OP50 to analyze only Sal, no PiPEG, remove_selected


```{r}

data_no_OP50 <- data_no_LBPiPEG[data_no_LBPiPEG$Strain != "OP50", ]

# fit
glm_no_OP50 <- glm(
  TD50 ~ Environment * Media,
  family = Gamma(link = "log"),
  data = data_no_OP50
)
summary(glm_no_OP50)

# Diagnostic
par(mfrow = c(2, 2))
plot(glm_no_OP50)

#Anova
library(car)
type_iii_anova_no_OP50 <- Anova(glm_no_OP50, type = "III")
print(type_iii_anova_no_OP50)

anova(glm_no_OP50)

library(ggplot2)
ggplot(data_no_OP50, aes(x = Media, y = TD50, fill = Environment)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "TD50 by Environment and Media (Excluding OP50)",
       x = "Media", y = "TD50")

```








```{r}



```








```{r}



```








```{r}



```








```{r}



```








```{r}



```








```{r}



```








```{r}



```








```{r}



```








```{r}



```






